### 2015-08-24 paper.tex
Initial import.

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\usepackage{makeidx}  % allows for indexgeneration
\usepackage{amssymb, amsmath, amsfonts, graphics, graphicx}
\usepackage{enumerate}

\usepackage{longtable}

\usepackage{url}
\urldef{\mailsa}\path|{alfred.hofmann, ursula.barth, ingrid.haas, frank.holzwarth,|
\urldef{\mailsb}\path|anna.kramer, leonie.kunz, christine.reiss, nicole.sator,|
\urldef{\mailsc}\path|erika.siebert-cole, peter.strasser, lncs}@springer.com|
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

%=======================================================================
%==========================DOCUMENT=====================================
%=======================================================================
\begin{document}

\newcommand{\fictAquantor}{\ensuremath{\forall\colon\boldsymbol{True}}}
\newcommand{\fictEquantor}{\ensuremath{\exists\colon\boldsymbol{True}}}
\newcommand{\bomega}{\boldsymbol{\omega}}
\newcommand{\bphi}{\boldsymbol{\phi}}
\newcommand{\eqdef}{\stackrel{\mathrm{df}}{=}}
\newcommand{\bigand}[2]{\raisebox{-4pt}{\ensuremath{\overset{#1}{\underset{#2}{\text{\huge\&\normalfont}}}}}}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Automated Theorem Proving Based on the Calculus of Positively Constructed Formulas}
% a short form should be given in case it is too long for the running head
\titlerunning{ATP Based on the PCF Calculus}

\author{Alexander Larionov\thanks{The work was supported by Irkutsk State University, grant No 2012-01-02} \and Ivan Terehin\thanks{The work was supported by Irkutsk State University, grant No 2012-01-01} \and Evgeny Cherkashin \and Artem Davydov
%\thanks{}
}

%\authorrunning{Lecture Notes in Computer Science: Authors' Instructions}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{Irkutsk State University,\\ Irkutsk, Karl Marks str. 1, Russia\\
Institute of System Dynamics and Control Theory SB RAS,\\ Lermontov str. 134, Irkutsk, Russia\\
%\mailsa\\
%\mailsb\\
%\mailsc\\
%\url{http://www.springer.com/lncs}
}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

%\toctitle{Lecture Notes in Computer Science}
%\tocauthor{Authors' Instructions}
\maketitle

%=======================================================================
%==========================ABSTRACT=====================================
%=======================================================================
\begin{abstract}
A description of the calculus of positively constructed formulas (PCF) and prover based on this calculus is considered in this paper. The PCF calculus has been developed by Russian scientists S.N. Vassiliev and A.K. Zherlov by an evolutionary way in describing and solving control theory problems. This calculus has features, which are applicable in control theory. The described implementation of the prover uses several techniques and strategies to improve prover performance. The prover is being tested by means of solving problems from TPTP library. The usage of implemented strategies is also commented in this paper.
\keywords{positively constructed formulas, automated theorem proving, proof strategies}
\end{abstract}

%=======================================================================
%==========================INTRODUCTION=================================
%=======================================================================
\section{Introduction}

Originally \cite{SNV1990,ICDS2000} the calculus of positively constructed formulas (PCF) was developed by Russian scientists S.N. Vassilyev and A.K. Zherlov by an evolutionary way in describing and solving control theory (CT) problems. In \cite{ICDS2000} the PCF calculus is presented as first-order logical formalism, examples of CT problems described  and solved by the PCF calculus (elevator group control, mobile robot action planning and telescope guidance), proof of soundness and completeness are presented as well.

The PCF calculus is positioned as not only machine-oriented, but also as a human-oriented, mostly aimed at solving the problems of dynamic systems control due to the following features of this calculus: unique inference rule and simple scheme of axioms; modifiability of semantics (constructive, monotonic, temporal, etc.) and besides it is possible to construct intuitionistic inferences of some non-Horn formulas; the explicit usage of $\forall$ and $\exists$ quantifiers, the scolemization procedure is not required.

However, in this paper, we will not cover the human--orientability properties of the PCF calculus and its dynamic systems control properties as well. We deal with the automatic search of a logical inference, as the inference engine capabilities are the basis of the problem solving systems, described in \cite{ICDS2000}. In order to estimate the applicability of the PCF calculus for automatic theorem proving we develop a prover program and test it on problems from TPTP library.

%This paper contains a description of the PCF calculus, an implementation of the prover program, and strategies of logical inference used to direct the inference search algorithms. The results of a comparison of the developed prover system with other provers are presented.

%=======================================================================
%==========================BACKGROUND===================================
%=======================================================================
\section{Background}

%As a theoretical basis for our prover we have used the PCF calculus. Let's describe it in more details.

The PCF calculus $\boldsymbol{JF}$ is a triple $\left\langle \boldsymbol{LF}, Ax\boldsymbol{JF}, \bomega \right\rangle$, where $\boldsymbol{LF}$ is the PCF language, $Ax\boldsymbol{JF}$ is the only axiom schema and $\bomega$ is the unique inference rule of $\boldsymbol{JF}.$

%language
\subsection{The Language of Positively Constructed Formulas}

Let us denote the set of all conjuncts as $Con$ and assume that a {\em conjunct} is either a finite set of usual atoms of a first-order language, or $\boldsymbol{False}$, where $\boldsymbol{False}$ satisfies the property $A \subset \boldsymbol{False} $ for any $A \in Con$. The empty conjunct is denoted by $\boldsymbol{True}$. It is clear that if $A \in Con$ then $A \cup \boldsymbol{False} = \boldsymbol{False}$. Atoms of any conjunct (except $\boldsymbol{True}$ and $\boldsymbol{False}$) may contain variables, constants and function symbols.

\begin{definition}\label{def:1}
Let $\bar{x}$ be a set of variables (possible empty set) and $A$ be a conjunct. Well-formed formulas of the PCF language are defined as follows:

An expression $\exists \bar{x}\colon A$ is a $\exists$-formula; an expression $\forall \bar{x}\colon A$ is a $\forall$-formula.

Let $G_1,\ldots,G_k$ be a $\exists$-formulas, then $\forall$-formula is a formula of the following form: $\forall \bar{x}\colon A\left(G_1,\ldots,G_k\right)$.

Let $G_1,\ldots,G_k$ be a $\forall$-formulas, then $\exists$-formula is a formula of the following form: $\exists \bar{x}\colon A\left(G_1,\ldots,G_k\right)$.

\end{definition}

An expression is a well--formed PCF if it is constructed based on above Definition \ref{def:1} rules only.

The variables from $\bar{x}$ are bound by a corresponding quantifiers, and these variables are called $\forall$-variables and $\exists$--variables correspondingly.

%Any $\forall$--variable that does not occur in the corresponding conjunct is called {\em unconfined} variable.

%semantic
Let us define a semantic of PCFs by a common semantics of a corresponding first-order formula (FOF) in the classical first--order logic (FOL).

\begin{definition}\label{def:semantic}
Let $A = \{A_1,\ldots,A_l\}$ be a conjunct and $\bar{x} = \{x_1,\ldots,x_n\}$ be a set of variables. By $A^{\&}$ we denote $A_1 \&\ldots\&A_l$ and $\boldsymbol{False}^{\&}= False, \boldsymbol{True}^{\&}=True$ (propositional constants). By $F^{\text{\tiny{FOF}}}$ we denote an image of PCF formula $F$ in the FOL.

%возможно убрать
If $F= \exists \bar{x}\colon A$ then $F^{\text{\tiny{FOF}}} = \exists x_1 \ldots \exists x_n \left(A^{\&}\right).$
%возможно убрать

If $F = \forall \bar{x}\colon A$ then $F^{\text{\tiny{FOF}}} = \forall x_1 \ldots \forall x_n \left(A^{\&}\right).$

If $F = \exists \bar{x}\colon A\left(G_1,\ldots,G_k\right)$ then $$F^{\text{\tiny{FOF}}} = \exists x_1 \ldots \exists x_n  \left(A^{\&} \& \left(G_{1}^{\text{\tiny{FOF}}} \& \ldots \& G_{k}^{\text{\tiny{FOF}}}\right)\right).$$

If $F = \forall \bar{x}\colon A\left(G_1,\ldots,G_k\right)$ then $$F^{\text{\tiny{FOF}}} = \forall x_1 \ldots \forall x_n \left(A^{\&}\rightarrow \left(G_{1}^{\text{\tiny{FOF}}} \vee\ldots\vee G_{k}^{\text{\tiny{FOF}}}\right)\right).$$

\end{definition}

Any PCF can be obviously represented by a tree structure. Each node of that tree has the following form $\exists \bar{x}\colon A$ or $\forall \bar{x}\colon A$. Therefore, for readability we will represent PCF as a tree, as well as we will use corresponding terminology: node, branch, root, leaf, etc.
%Каждый узел дерева имеет вид или $\forall\bar{x}\colon A$ or $\exists\bar{x}\colon A$. Например, в формуле вида $\forall \bar{x}\colon A\left(G_1,\ldots,G_k\right)$ корневым узлом являюется узел $\forall \bar{x}\colon A$, а его дочерние узлы это $G_1,\ldots,G_k\right$.

If a PCF $F$ roots from the $\forall\colon\boldsymbol{True}$ node and each leaf of $F$ is a $\exists$-node then $F$ is called a PCF in the {\em canonical form}.
It is clear, that any PCF $F$ can be converted to the canonical form by the following transformations:
\begin{enumerate}
\item If $F$ is not canonical $\forall$-formula then $\forall\colon \boldsymbol{True}\left(\exists\colon \boldsymbol{True}\left(F\right)\right)$ is PCF rooted from the $\forall\colon\boldsymbol{True}$.
\item If $F$ is $\exists$-formula then $\forall\colon \boldsymbol{True}\left(F\right)$ is PCF rooted from the $\forall\colon\boldsymbol{True}$.
\item If $F$ has leaf $\forall \bar{x}\colon A$ then the node $\exists\colon\boldsymbol{False}$ can be added as child of this leaf.
\end{enumerate}

%как правило
In the sequel, only PCFs in the canonical form are considered.

%ПО-формулы имеют древовидную структуру, а значит можно использовать соответствующую терминологию и изображать их как деревья.+++++++++++++++++++++++++
%PCFs have a tree-like structure, therefore we can use corresponding terminology and represent PCFs as trees.
%Any $\forall$--variable that does not occur in the corresponding conjunct is called {\em unconfined} variable.

Some parts of a PCF are denoted by particular terms: the topmost (0-depth) node is called {\em root}; a 1-depth node is called {\em base} (of formula); a maximal subtree that rooted at the 1-depth node is called {\em base subformula}; a 2-depth node is called {\em question} (to its base); a maximal subtree that rooted at the 2-depth node is called {\em question subformula}; a maximal subtree that rooted at 3-depth node is called {\em consequent} (subformula). With respect to Definition~\ref{def:semantic} if an even-depth (odd-depth) node has more than one children then the node has {\em disjunctive branching} (correspondingly {\em conjunctive branching}).%; if a node of PCF has a following form $\forall \bar{x}\colon A$ and if $\exists v \in \bar{x}$ and $v \notin A$ then variable $v$ is called {\em unconfined} variable. %The semantics of the denotation used results from the above mentioned human--orientation of the PCF calculi.

\begin{example}
Let us consider a formula of FOL.
$$F= \neg\bigl(\forall x\:\exists y P(x,y)\rightarrow \exists z P(z,z)\bigr).$$
An image $F^{\text{\tiny{PCF}}}$ of $F$ in the PCF language is

$$F^{\text{\tiny{PCF}}} = \forall\colon \boldsymbol{True}(\exists\colon\boldsymbol{True}(\forall x\colon\boldsymbol{True}(\exists y\colon P(x,y)), \forall z\colon P(z,z)(\exists\colon\boldsymbol{False}))).$$
%А эта ПО-формула в древовидной форме имеет следующий вид+++++++++++++++++++++++++
The tree--like form of the above PCF is as follows:
$$F^{\text{\tiny{PCF}}} = \forall\colon \boldsymbol{True}-\exists\colon\boldsymbol{True} \left\{
\begin{array}{lcl}
 \forall x\colon\boldsymbol{True} & - & \exists y\colon P(x,y) \\
 \forall z\colon P(z,z) & - & \exists\colon\boldsymbol{False}.
\end{array}
\right.$$
\end{example}
%Для удобствоа мы используем символ --- для обозначения рёбер в древовидном представлении ПО-формулы. Фугрная скобка также является вспомогательным символом.+++++++++++++++++++++++++
For convenience we use symbols ``--'' and ``\{'' for denoting the edges and branching in tree--like representation of PCF.

%--------------calculus
\subsection{The Calculus of Positively Constructed Formulas}

The axiom schema $Ax\boldsymbol{JF}$ of the PCF calculus $\boldsymbol{JF}$ has the following form:
$$\forall\colon\boldsymbol{True}\left(\exists \bar{x}_1\colon\boldsymbol{False}\left(\widetilde{\Phi}_1\right),\ldots,\exists \bar{x}_n\colon\boldsymbol{False}\left(\widetilde{\Phi}_n\right)\right). $$

In the PCF calculus to prove $F$ we attempt to refute its negation, that's why the axioms are identical to false. Thus, an inference process in the PCF calculus is a {\em refutation} process.

\begin{definition}
\label{ircond}
Let us assume that a question $\forall \bar{y}\colon A$ to a base $\exists \bar{x}\colon B$ has an answer $\theta$  if and only if $\theta$ is a substitution $\bar{y} \rightarrow H^{\infty}$ and $A\theta \subseteq B$, where $H^{\infty}$ is Herbrand universum based on variables from $\bar{x}$, constants and function symbols that occur in corresponding base subformula.
\end{definition}

%The unique unary inference rule $\bomega$ is defined as follows.

\begin{definition}
If a PCF $F$ has the structure $\forall\colon\boldsymbol{True}\left(\exists \bar{x}\colon B\left(\Phi\right),\Sigma\right)$, where $\Sigma$ is a list of other base subformulas, and $\Phi$ is a list of question subformulas, that contains a question subformula $\forall \bar{y}\colon A(\exists \bar{z}_i\colon C_i\left(\Psi_i\right))_{i=\overline{1,k}}$, then a formula $\bomega F$ is an application result of the unary inference rule $\bomega$ to the question $\forall \bar{y}\colon A$ with an answer $\theta$ and $\bomega F = \forall\colon\boldsymbol{True}(\exists \bar{x} \cup \bar{z}_i\colon B \cup C_i\theta\left(\Phi \cup \Psi_i\theta\right)_{i=\overline{1,k}},\Sigma).$
\end{definition}

After appropriate renaming some bound variables inside each subformula the expression $\bomega F$ will satisfy the well--formed PCF requirements. %in the canonical form requirements

%
Any finite sequence of PCFs $F, \bomega F, \bomega^2 F,\ldots,\bomega^n F$, where $\bomega^n F \in Ax\boldsymbol{JF}$, is called a {\em refutation} of $F$ in the PCF calculus.

A question with $\exists:\boldsymbol{False}$ consequent node is called {\em goal question}. If a question has disjunctive branching then answering this question results in formula {\em splitting}.

For clarity consider simple example of a PCF refutation.

\begin{example}[A refutation in $\boldsymbol{JF}$]\label{proofexample}

\begin{equation*}\label{ex:f1}
  F_1 = \forall\colon\boldsymbol{True} - \exists\colon S(e)(Q_1,Q_2,Q_3,Q_4);
\end{equation*}
\begin{equation*}
  \begin{array}{l}
  Q_1 = \forall x\colon S(x) - \exists\colon A(a); \\
  Q_2 = \forall x,y\colon C(x),D(y) - \exists\colon\boldsymbol{False}; \\
  Q_3 = \forall x,y\colon B(x),C(f(y)) - \exists\colon\boldsymbol{False}; \\
  Q_4 =
  \forall x\colon A(x) -
  \left\lbrace
  \begin{array}{l}
    \exists y\colon B(y),C(f(x)) \\
    \exists \colon C(x) - \forall z\colon A(z),C(z) - \exists\colon D(f(z)).
  \end{array}\right.
  \end{array}
\end{equation*}
At the first step only one answer $\{x \rightarrow e\}$ to question $Q_1$ exists. After applying rule $\bomega$ with this answer, formula will be in the following form:
\begin{equation*}\label{ex:f2}
  F_2 = \forall\colon\boldsymbol{True} - \exists\colon S(e),A(a)(Q_1,Q_2,Q_3,Q_4).
\end{equation*}
At the second step also only one answer $\{x \rightarrow a\}$ to  question $Q_4$ exists. After applying $\bomega$ with this answer, formula is split, because $Q_4$ has disjunctive branching. And now formula has the following form:

\begin{equation*}\label{ex:f3}
F_3 =
\forall:\boldsymbol{True} -
\left\lbrace
\begin{array}{l}
  \exists y_1\colon S(e),A(a),B(y_1),C(f(a)) -
  \left\lbrace
  \begin{array}{l}
    Q_1 \\ \cdots \\ Q_4
  \end{array}\right. \\
  \exists\colon S(e)A(a),C(a) -
  \left\lbrace
  \begin{array}{l}
    Q_1 \\ \cdots \\ Q_4 \\
    \forall z\colon A(z),C(z) - \exists\colon D(f(z)).
  \end{array}\right. \\
\end{array}\right.
\end{equation*}

At the third step first base can be refuted by answering on $Q_3$ goal question with $\{x \rightarrow y_1; y \rightarrow a\}$. Refuted base (base subformula) can be deleted from the list of base subformulas.

At the fourth step the answer $\{z \rightarrow a\}$ to fifth new question exists. And now formula has the following form:
\begin{equation*}\label{ex:f5}
  F_4 = \forall\colon\boldsymbol{True} - \exists\colon S(e),A(a), C(a),D(f(a))
  \left\lbrace
  \begin{array}{l}
    Q_1 \\ \cdots \\ Q_4 \\
    \forall z\colon A(z),C(z) - \exists\colon D(f(z)).
  \end{array}\right.
\end{equation*}

At the fifth step the only base can be refuted by answering on $Q_4$ goal question with  answer $\{x \rightarrow a; y \rightarrow f(a)\}$.

The refutation is finished because all bases were refuted.

\end{example}

%=======================================================================
%==========================TECHNICQUES==================================
%=======================================================================
\section{Prover implementation}

\subsection{Implementation language}

Our prover is implemented in the D programming language \cite{DPL1,DPL2}. The D programming language has many features: compiler produce a native binary code, which both fast and optimized; procedures of memory management are manually adjustable, but it is also possible to use garbage collection; there is a good library for parallel computation construction in Erlang--like model, which is generally recognized as the quite good \cite{Erlang1}.

\subsection{Proof State Tree}

%---------
%One of the basic components of our prover implementation is a proof states tree (PST).

Each step of the inference in the PCF calculus can be considered as an addition of corresponding consequent subformulas to the base subformula (facts to a base; questions to a question lists). Thus, a formula is usually monotonically increasing, and splitting if a disjunctive branching happens.
%---------

One of the main tools of the logical inference search implementation in the developed prover is the \emph{proof state tree} (PST), which is built with the data structures based on so--called chunks described below. The PST structures are used to store the sequence of the steps of logical inference. Each step corresponds to a single operation that were performed. PST also represents the current state of the formula under refutation. The main purpose of PST is to strictly fix all operations and their parameters, which occurs on every step of logical inference construction. The example of such operation is the fact of applying a substitution in a base subformula to a question.

%--------------------------------------------------------------------------------
\subsubsection{Chunk.}
Chunk is a single--linked list of elements with type {\tt T} which have not only reference to the \texttt{first} and \texttt{last} elements of that list. A chunk is empty, if it does not contain any elements, and \texttt{first} and \texttt{last} elements are NULL--references. We say, that chunk $C_1$ is \emph{linked} with chunk $C_2$ \emph {from the left}, if the \texttt{last} node of chunk $C_1$ is referring to the \texttt{first} node of chunk $C_2$ and, simultaneously, $C_2$ linked with $C_1$ \emph{from the right}. For this operation a procedure of linking {\tt void link(Chunk!(T) c)} is defined, which links current chunk and chunk {\tt c} from left. If the chunks are not empty, then linking is proceeded according to the definition. If one of the chunks is empty, then its \texttt{first} and \texttt{last} elements will be assigned to the value of the linking element. Note, that several chunks can be linked with different chunk from one of its sides simultaneously. Therefore, using the procedure of linking one can build a tree of chunks, which grows from leafs.

%With the chunks we can organize main storage of information of PCF and the states of logical inference.

In Fig.~\ref{fig:chunk1} an example of structure is presented, which is built from chunks.
\begin{figure}[h]
  \vspace{0.5cm}
  \centering
  \includegraphics[width=0.6\linewidth]{img/ChunkFullp1.eps}
  \caption{Chunks}
  \label{fig:chunk1}
\end{figure}
There are five chunks in this figure, and several of them are linked together with \texttt{first} field: ``5'' and ``4'', ``4'' and ``3'', ``3'' and ``1'', ``2'' and ``1''. Chunk 4 is empty. \texttt{first} and \texttt{last} is the first and last nodes of chunks. Numbers are the identifiers of the corresponding nodes. Empty chunk duplicates node \texttt{first} for the node linked with empty chunk from the left.

%---------------------
\subsubsection{Proof state tree.}
A proof state tree is a tree with the following properties: the root is the original PCF, other nodes are consequent subformulas (with the answer applied to them and variables renamed). In addition, the nodes contain some system information like step of inference, answer, etc. Thus, the number of the leafs is the current number of the bases, and at current inference step a PCF is denoted by the path from the leaf to the PST root. This approach allows us to rollback (backtrack) the inference search process, observe proof states and share references among terms and formulas. If a base is refuted then its memory can be released by deleting of the path from corresponding leaf to the soonest branching.

Every node of PST is a chunk. The growth of the tree is a linking of new chunks with the existing leafs of the tree from the left.

PST for the formula from Example \ref{proofexample} is represented in Fig.~\ref{fig:pst}.
PST root is an original PCF $F_1$. Node ``2'' is a  $Q_1$ question consequent $\exists\colon A(a)$, and the path from node ``2'' to the root node corresponds to PCF $F_2$. Nodes ``3'' and ``4'' are corresponding consequents for $Q_4$. The path from node ``3'' to the root node and the path from node ``4'' to the root node correspond to base subformulas for PCF $F_3$. For example, formulas that are denoted by paths ``5''--``1'' and ``3''--``1'' share data that are represented by nodes ``1''--``2''. When base subformula that is represented by path ``3''--``1'' is refuted, we can delete path from node ``3'' to soonest branching --- in this case only node ``3'' is deleted, because nodes ``2''--``1'' are still used for representing another base subformulas.
%The lettering in figures should have a height of 2mm (10-point type)
%caption UNDER the figures
\begin{figure}[h]
  %\vspace{0.5cm}
  \centering
  \includegraphics[width=0.4\linewidth]{img/PSTp1.eps}
  \caption{Proof state tree for PCF from Example \ref{proofexample}.}
  \label{fig:pst}
\end{figure}

%---------------------
\subsubsection{Data sharing: sharing of base subformulas.}
%------
In addition, in a PST the data--sharing (memory sharing) is implemented: if two base subformulas are characterized by the corresponding paths, their bases share data, which are represented by the same part of these paths.
\begin{figure}[h]
  \vspace{0.5cm}
  \centering
  \includegraphics[width=0.4\linewidth]{img/Graphsp1.eps}
  \caption{Example of proof state tree for base subformula refutation}
  \label{fig:datasharing1}
\end{figure}

If a PCF contains a question with disjunctive branching, then its base subformula will be divided into several new base subformulas during the searching procedure of the logical inference. The number of new subformulas corresponds to the number of disjunctive subformulas in the consequent of the question. In the straightforward variant of the prover implementation, the sharing will require copying previous state of the original formula more than once; this copying have a linear complexity, but it leads to the redundant expenses of memory and processor time. PST allows one to implement sharing the common subbranches. Let's we have the following base subformula:
$$\exists: A(a) - \forall x: A(x) - \left\{
\begin{array}{lcl}
 \exists \colon B(x) & - & \forall y: B(y) - \exists\colon\boldsymbol{False}\\
 \exists \colon C(x) & - & \forall y: C(y) - \exists\colon\boldsymbol{False}
\end{array}
\right. $$

Note, that the formula has disjunctive branching and it is deep. PST for the refutation of this base subformula is presented on Fig.~\ref{fig:datasharing2}
\begin{figure}[h]
  \vspace{0.5cm}
  \centering
  \includegraphics[width=0.82\linewidth]{img/DataSharing2p1.eps}
  \caption{Example of proof state tree for base subformula refutation}
  \label{fig:datasharing2}
\end{figure}

%====================================================================================
%====================================================================================
\subsection{Lazy concretization}

By the Definition~3 the inference rule $\omega$ can be applied if the question $\forall \bar{y}\colon A$ to base $\exists \bar{x}\colon B$ has an answer $\theta$, such that $\theta$ is a substitution $\bar{y} \rightarrow H^{\infty}$ and $A\theta \subseteq B$. If all the variables in the $\bar{y}$ are occurred in $A$, then the search of corresponding elements from $H^{\infty}$ is carried on by the algorithm of subsumption and depends on only corresponding base of facts, hence the conditions of sets subsumptions need to be satisfied. If a variable in the $\bar{y}$ is not occurred in $A$ then the variable is called {\emph unconfined variable} and it is not used in a subsumption. However, we need to find a substitution for these variables as well, and any element of Herbrand universum is a correct substitution. The problem with Herbrand universum is that in the general case, it is an unlimited set and it is not known what element should be taken as a substitution.

The main concept of lazy concretization is as follows. Instead of selection of a concrete element of the Herbrand universum for a substitution of an unconfined variable an {\em unspecified Herbrand element} (UHE) is used temporally. The UHE will be concretized further to a proper term. Such concretization is based on the logical inference requirements. UHE will be concretized to a ground term, or in some situations, it will remain unconcretized. In such case, the required terms provide the possibility of answering to the question: UHE is gradually concretized in such way, that on every step of inference we can build new answer for some question. Simultaneously the following problem is being solved: to what term we need to concretize UHE to make it possible to answer the question? Note, that originally UHE appears as a part of answer in one question and its concretization is done while searching the answers for other questions. As ``gradually concretization'' we mean procedure of lazy computations: UHE concretized as precise, as it is enough for answering the current question, i.e. concretization, generally, is partial, for example UHE for $h$ is concretized to $f(h_1)$, where $h_1$ is a new UHE.

By it's nature UHE is similar to unconcretized $\forall$-variable in the sense that it is mutable, i.e. it is potentially concretized, however all such changes must be directed only to concretization of UHE, i.e. UHE is substituted only to some term or to different UHE, but not to the variable.

Let's consider the example of this technique in process of building the logical inference for the following formula.
\begin{equation*}
  \forall\colon\boldsymbol{True} - \exists\colon\boldsymbol{True} -
  \left\lbrace
  \begin{array}{l}
    \forall x\colon\boldsymbol{True} - \exists\colon A(x) \\
    \forall x\colon A(f(x)) - \exists\colon B(f(x)) \\
    \forall\colon B(f(a)) - \exists\colon \boldsymbol{False}
  \end{array}\right.
\end{equation*}

At the first step an answer $\{x \rightarrow h_1\}$ to the first question exists, $x$ is an unconfined variable and $h_1$ is an unspecified Herbrand element. After this step the atom $A(h_1)$ is added to the base. At the second step an answer $\{x \rightarrow h_2\}$ to the second question exists and $h_1$ is concretized to $f(h_2)$, after this step $B(f(h_2))$ is added to the base. Finally, at the third step the trivial (empty) answer exists to the third question and $h_2$ is concretized to $a$.

For efficiency increase in this strategy, we are using two constraints.

%Для стртегии ленивой конкретизации предложено два типа ограничений
\subsubsection{Constraint~1: Limit of concretizations number.}
%Ограничение количества конкретизаций.

%---
Two UHEs are \emph{similar}, if they appear as the result of a substitution for the same unconfined variable. This situation appears if for a question with unconfined variables a sequence of answers were constructed, such that on every step of the sequence new UHE created as a substitution. A constraint for the similar UHE--concretizations excludes such situations, when UHE is concretized on a step, where it is not necessary, and, consequently, an opportunity to concretize it appears on other step. For every UHE a reference to the unconfined variable is stored, which the UHE was selected as a substitution for. The set of terms, which concretized a unconfined variable through its UHE, is stored in the corresponding question. The engine allows tracking the UHE concretization. If a predefined limit of the concretizations is exceeded then all further concretizations are denied. The limit is chosen by user using additional information about the problem under investigation, or it is taken as $1$, and increased after an unsuccessfull logical inference, then the inference attempt is repeated.

\subsubsection{Constraint~2: Storing expressions which contain UHE.}
%Сохранение выражений, содержащих НЭЭ

For a question with unconfined variables, one can apply all available answers, i.e. with using all elements of Herbrand universum. It is known, that this is not technically possible because of unlimited Herbrand universum. So, the technique of lazy computations is used. In ideal case, all answers for question is used simultaneously, but actually only the required questions are used. For example, if in a some formula $\forall x: True  - \exists A(x)$, and $H^{\infty}= \{a, f(a), f(f(a)), ...\}$ then usage of all available questions simultaneously leads to addition of $\{A(a), A(f(a)), A(f(f(a))), ...\}$ elements (amount is unlimited) to the base. Instead, this set is associated with the element $A(h)$, where $h$ is UHE, and this element generate required elements of $H^{\infty}$ set in further.

In such case, one can find a similarity with lazy concretization strategy without constraints. The advantage of the approach with storing expressions, which contain UHE, is that the expression with UHE is always stored, i.e. it is not possible to lose it by applying some substitutions.

%====================================================================================
%-----------------k,m,-condition-----------------
\subsection{Strategy ``k,m-condition"}
An answer to a question is accepted if at some point of the following $k$ steps of its inference a certain condition becomes true at least $m$ times. This strategy is used in the following cases.

\subsubsection{$k,m$--refutation.} An answer to a question with disjunctive branching is accepted if in next $k$ steps at least $m$ generated bases were refuted, otherwise the process restarted with a new answer. This strategy allows us to constrain highly spawning inferences by means of control answering questions with disjunctive branching.

\subsubsection{$k$--unrefutation.}
For a question with disjunctive branching an answer is accepted if during next $k$ steps no refutation for the base subformula was achieved, using the questions without disjunctive branching. This strategy, as it is in the previous variant, reduces growth of a formula due to answering questions with disjunctive branching. If no refutation can be constructed using this strategy, then answering to the questions allowed. The parameter $m$ is dropped in that case.

\subsubsection{$k,m$--concretization.}
An answer for a question is accepted if during next $k$ steps at least $m$ UHE are concretized. This strategy is also aimed at limiting a complexity of the formula. Large amounts of additional information and conditions are associated with unconcretized UHE. After an UHE is concretized, the associations are released. So as soon we process the UHE as soon we will take advantage of additional memory.

%====================================================================================
%Parallel strategy
\subsection{Parallel strategy}

Let's consider techniques and strategies for parallel implementations of logical inference algorithms in the PCF calculus.

\subsubsection{First parallel strategy.}
If a question has disjunctive branching then, after we have answered that question, the formula will splits and transforms to the formula with number of bases equals to the number of answered question consequents, i.e. in general case the number of the base subformulas is increasing on every step of logical inference. In addition, the original formalization of a problem in PCF language can contain more than one base subformula. In order to refute the original PCF formula every base need to be refuted. The refutation of the base subformulas can be processed in parallel, i.e., the refutation of each base is processed in an independent thread. This is possible if the bases are independent from each other, i.e. they don't share $\forall$--variables and UHEs. Bases never share $\forall$--variables because bases don't contain $\forall$--variables at all (by definitions), as they contain only ground terms. However, bases can share the UHEs in case of usage of the lazy concretization strategy. Therefore, this strategy can be applied in case if the lazy concretization strategy is not used. According to the above described features of PCF we say, that PCF have property called ``OR--parallelism''.

\subsubsection{Second parallel strategy.}
In addition to the OR--parallelism, parallel schemes of algorithms for searching answers for each questions can be constructed. Since the questions do not share variables, as in previous strategy, the answer search procedure can be performed also in an individual computation thread. In addition, like in the previous strategy, second parallel strategy formulated as follows: each answer search procedure to a question is independent of other questions.

\subsubsection{Recomendation.}
One recommendation for an implementation of the described algorithms on cluster computing systems can be given: there is a problem to divide tasks between computing nodes depending on connection speed between the nodes. For example, implementation of the first strategy must bind process to the computing node. Binding the process to the computing node in case of second strategy is possible with increasing the conjunct of question and increasing the set of substitutions corresponding to the every atom of question. Otherwise, communication expenses can override the time of computation and will degrade the overall result.

\subsection{Memory manager}
Allocation of memory for fixed size records is implemented based on standard approach, which is similar to \cite{gmemory}. Array of structures is allocated with memory manager of operation system and then converted to freelists. New structures are allocated from the freelists, and released ones are returned back to the freelists. If a list is exhausted a new array is allocated from heap memory.

In the prover for each variant of records, own array is allocated and own freelist constructed. The garbage collection is synchronized with garbage collector of the D programming language runtime engine.

%=======================================================================
%==========================EXPERIMENTS==================================
%=======================================================================
\section{Experiments}
Testing the developed prover was performed on problems form TPTP \cite{tptp} library version 5.4.0 (Thousands of Problems for Theorem Provers, www.tptp.org). This library is the de facto standard for testing the ATP provers. The type of problems that were chosen are the problems of FOF without equality. Total number of problems is 1221 with rating from 0.0 to 1.0. Total number of solved problems is 780. The maximal rating of solved problem is 0.62. The best results among domains presented in the Table~1.

\begin{table}
\caption{Domains}
\begin{tabular}{|p{0.3\linewidth}|p{0.4\linewidth}|p{0.1\linewidth}|}

\hline
\textbf{Domain} & \textbf{No of problems in the domain} & \textbf{Solved} \\
\hline
Geometry (GEO) & 242 & 204 \\
\hline
Management (MGT) & 22 & 22 \\
\hline
Syntactic  (SYN) & 275 & 180 \\
\hline
Semantic Web  (SWB) & 25 & 22 \\
\hline
\end{tabular}
\end{table}

10 hardest problems solved by our prover is presented in the Table~2.
\begin{table}
\caption{Hardest problems solved by our prover}
\begin{tabular}{|p{0.2\linewidth}|p{0.2\linewidth}|p{0.2\linewidth}|p{0.2\linewidth}|}

\hline
\textbf{Problem} & \textbf{Rating} & \textbf{Time (s)} & \textbf{Steps number} \\
\hline
COM008+1 & 0.62 & 1.9 & 33551 \\
\hline
LCL640+1.005 &  0.62 &  0.06 &  821 \\
\hline
LCL656+1.010 &  0.58 &  1.5 &  43 \\
\hline
SWB012+3 &  0.54 &  65.64 &  102428 \\
\hline
SYN353+1 &  0.54 &  0.002 &  33 \\
\hline
SYN548+1 &  0.54 &  0.01 &  15 \\
\hline
SWB020+2 &  0.5 &  0.6 &  618 \\
\hline
LCL666+1.005 &  0.5 &  0.35 &  2592 \\
\hline
KRS235+1 & 0.46 & 5.4 & 24406 \\
\hline
SWB029+3 &  0.38 &  55.27 &  98471 \\
\hline
\end{tabular}
\end{table}

%График...

Prover shown good results for problems, whose formalization in the PCF language contains quite large conjuncts.

\subsection{Translation from TPTP to PCF}
TPTP contains an axiom library, which consists of different files. Formula to be investigated is collected from the files. The collection is performed with special utilities such as \texttt{tptp4X}. The collected subformulas are passed to a translator from TPTP representation to the input language of the prover. The translator is implemented with the \texttt{hotptp-yl-parser-verbose} \cite{TPTPTrans} utility that translates input TPTP--files to abstract syntax trees. Translation component of the utility is generated on the base of syntactic analysis of the current specification of TPTP. Therefore, in each case of the specification refinement, the utility can be automatically updated (recompiled). Using this approach, we are able to adopt our prover to the new input language specification.

The abstract syntax tree is passed to translation component that imports the tree, translates the predicate formula or CNF, performs some reduction of subformulas by applying a set of rules, converts the formulas into the PCF language, performs additional reduction of the PCFs, renames the quantifier variables, and, at the last step, outputs the conversion result as textual representation of the problem in the language of the prover.

\subsection{Example of Formula}
Problem SYN380+1.p from TPTP library in our prover input language has a following form:
\begin{quote}
\texttt{\raggedright\noindent
\{\\
~~e[][]~\{\\
~~~~a[W][big\_r(W,W)]~\{\\
~~~~~~e[][False]~\{\}\};\\
~~~~a[X,Y][]~\{\\
~~~~~~e[][big\_r(X,Y)]~\{\};\\
~~~~~~e[Z][big\_q(Y,X)]~\{\\
~~~~~~~~a[][big\_q(Z,Z)]~\{\\
~~~~~~~~~~e[][False]~\{\}\}\}\}\}\\
\}\\
}
\end{quote}

\subsection{Example of solved problem}
Let us consider COM003+1.p (``The halting problem is undecidable'') problem with 0.33 rating.

The results of solving this problem by our prover and other provers are presented in Table~3.

\begin{table}
\caption{COM003+1.p solving}
\begin{tabular}{|p{0.2\linewidth}|p{0.35\linewidth}|}

\hline
\textbf{Prover} & \textbf{Time (s)} \\
\hline
Our prover & 0.05 \\
\hline
Vampire & 0.01 \\
\hline
iProver & 0.27 \\
\hline
leanCoP & 19.25 \\
\hline
EP & 54.45  \\
\hline
Zenon & 0.08 \\
\hline
\end{tabular}
\end{table}

%Рассмотрим задачу TPTP SYN548+1.

%В языке ПО-формул формализация выглядит следующим образом:

%Рейтинг 0.54. Время решения: 5.8с. Количество шагов: 82764.

%При решении данной задачи использовались стратегии: ленивая конкретизация (ограничение 1), $8$--неопровержимость, $8,1$--конкретизация, разделение данных по базе, агрессивное разделение термов.

%Let's consider problem SYN548+1 from TPTP.
%In PCF language formalization looks like the following:

%Rating is 0.54. Time for solving 5.8s. Number of steps: 82764.

%The lazy concretization (first constraint), $8$--unrefutation, $8,1$--concretization, data sharing in bases.

%==========================================================
\subsection{Comments on strategies}

%=============================
\subsubsection{Lazy concretization.}
Lazy concretization strategy allows one in most cases to find a substitution for an unconfined variable in acceptable time as compared to the sequential enumeration of Herbrand universum. The constraints are the most valuable part of this strategy, which allow to decrease the number of steps of logical inference in comparison to the unconstrained strategy. The decrease of the number of steps leads to a decrease of the problem solving time. Application of the constraint~1 is the best way of the inference time reduction in the case when a formula has questions with disjunctive branching and the constraint~2 in case if the formula contains no questions without disjunctive branching.

%=============================
\subsubsection{$k,m$--condition.}
The efficiency of this strategy reveals in the problems with disjunctive branching. In all cases of acceptable application of this strategy and correctly chosen parameters, the noticeable improvement is achieved.
%===========================
\subsubsection{Parallel stratgies.}
The important quality of parallel schemes of algorithms is their scalability, i.e., the performance improvement respect to increase of the number of computational nodes (CNs). That’s why the main characteristic of the prover productivity is not the time of execution of the program, but the ratio of program execution time in parallel mode on a given number of CNs to the program execution time on a single CN. The ratio is accounted for various number of CNs.

Experiments were carried out on the logical problems which PCF--formalizations have necessary properties for the parallel strategies: disjunctive branching, large number of questions, large conjuncts in questions. In Fig.~\ref{fig:parallel} the results of experiments are shown.

\begin{figure}[h]
  %\vspace{0.5cm}
  \centering
  \includegraphics[width=0.8\linewidth]{img/Parallel.eps}
  \caption{Results of parallel strategies testing}
  \label{fig:parallel}
\end{figure}

%=======================================================================
%==========================CONCLUSION===================================
%=======================================================================
\section*{Conclusion}
In this paper, we presented the results of the PCF calculus investigation in the context of the automatic theorem proving. The main result is a prover program implemented based on the PCF calculus. Several logical inference strategies were devised, adopted from other papers and implemented in this prover: memory sharing, lazy concretization, $k,m$--condition, parallel strategies. A special global data structure (Proof state tree) has been developed. Most of the strategies are implemented on the base of the structure.

Testing our prover has been carried out on problems from TPTP library. Testing has shown that the PCF calculus is suitable for automated theorem proving and some problems are solved more efficiently than other provers. The classes of problems, which are the most suitable for our system, are described.

Further investigations aimed at implementation of a productive logical inference engine with equality and with indexing techniques, for example, substitution tree indexing \cite{subtree}. The application of this prover for solving the problems of dynamic systems control will be investigated also.

%=======================================================================
%==========================BIBLIOGRAPHY=================================
%=======================================================================
\begin{thebibliography}{4}

\bibitem{DPL2} Alexandrescu, A. The D Programming Language. -- Addison-Wesley Professional, 2010. -- 460p.

\bibitem{DPL1} Armstrong, J. D Programming Language. URL: http://www.digitalmars.com/d/.

\bibitem{Erlang1} Armstrong, J. Programming Erlang.  -- The Pragmatic Programmers, 2007. -- 519p.

\bibitem{TPTPTrans} Gelder, A.V. TPTPparser utility. 2006. URL: http://users.soe.ucsc.edu/\~{}avg/TPTPparser/.

\bibitem{gmemory} GLIB: Memory Slices. URL: http://developer.gnome.org/glib/2.34/glib-Memory-Slices.html.

\bibitem{subtree} Graf, P. Substitution Tree Indexing // Proceedings of the 6th International Conference on Rewriting Techniques and Applications. 1995. pp.~117--131.

%\bibitem{HAR} Robinson, A., Voronkov, A. (eds.) Handbook of Automated Reasoning. -- MIT Press, Cambridge, 2001. -- 2150p.

\bibitem{tptp} Sutcliffe, G. The TPTP Problem Library and Associated Infrastructure: The FOF and CNF Parts, v3.5.0 // Journal of Automated Reasoning. V43, N4, pp.337--362.

\bibitem{SNV1990} Vassilyev, S.N.: Machine Synthesis of Mathematical Theorems. The Journal of Logic programming, V.9, No.2--3, pp. 235--266 (1990)

\bibitem{ICDS2000} Vassilyev, S.N., Zherlov, A.K., Fedunov, E.A.,
Fedosov, B.E.: Intelligent Control of Dynamic Systems. Moscow: Fizmatlit (2000)(in Russian)

\end{thebibliography}

\end{document}

### 2015-08-24 paper.tex
Added beginning from a better paper.

@@
 
-\author{Alexander Larionov\thanks{The work was supported by Irkutsk State University, grant No 2012-01-02} \and Ivan Terehin\thanks{The work was supported by Irkutsk State University, grant No 2012-01-01} \and Evgeny Cherkashin \and Artem Davydov
-%\thanks{}
-}
-
+% the name(s) of the author(s) follow(s) next
+%
+% NB: Chinese authors should write their first names(s) in front of
+% their surnames. This ensures that the names appear correctly in
+% the running heads and the author index.
+%
+\author{Alexander Larionov%
+\thanks{Ok}%
+\and Artem Davydov\and Evgeny Cherkashin\and Ivan Terekhin}
+%
 %\authorrunning{Lecture Notes in Computer Science: Authors' Instructions}
@@
 % unless you accept that it will be published
-\institute{Irkutsk State University,\\ Irkutsk, Karl Marks str. 1, Russia\\
-Institute of System Dynamics and Control Theory SB RAS,\\ Lermontov str. 134, Irkutsk, Russia\\
+\institute{Irkutsk State University,
+              1, Karl Marks str., Irkutsk, Russia \\
+           Institute for System Dynamics and Control Theory,
+              134, Lermontov str., Irkutsk, Russia   
 %\mailsa\\
@@
 
-%\toctitle{Lecture Notes in Computer Science}
-%\tocauthor{Authors' Instructions}
+\toctitle{Lecture Notes in Computer Science}
+\tocauthor{Authors' Instructions}
 \maketitle
@@
 \section{Background}
+We consider the language of first-order logic consist in formulas build on atomic formulas with $\&, \vee, \neg, \rightarrow, \leftrightarrow$ operators, $\forall$ and $\exists$ quantifier symbols and constant operators $true$ and $false$.
+The terms term, atom, literal we define in the usual way.
+FOFs.
 
-%As a theoretical basis for our prover we have used the PCF calculus. Let's describe it in more details.
+Let $X = \{x_1,\ldots,x_k\}$ be the set of variables, $A = \{A_1,\ldots,A_m\}$ be the set of atomic formulas, and $F = \{F_1,\ldots,F_n\}$ be the set of some FOFs.
+Then the following formulas $((\forall x_1) \ldots (\forall x_k) (A_1 \& \ldots \& A_m \rightarrow (F_1 \vee \ldots \vee F_n)))$ and $((\exists x_1) \ldots (\exists x_k) (A_1 \& \ldots \& A_m \& (F_1 \& \ldots \& F_n)))$ are denoted as  $\forall_XA\colon F$ and $\exists_XA\colon F$ respectively, keeping in mind that the $\forall$--quantifier corresponds to $\rightarrow F$, where $F$ means disjunction of all formulas from $F$, and $\exists$--quantifier corresponds to $\& F$, where $F$ means conjunctions of all formulas from $F$.
 
-The PCF calculus $\boldsymbol{JF}$ is a triple $\left\langle \boldsymbol{LF}, Ax\boldsymbol{JF}, \bomega \right\rangle$, where $\boldsymbol{LF}$ is the PCF language, $Ax\boldsymbol{JF}$ is the only axiom schema and $\bomega$ is the unique inference rule of $\boldsymbol{JF}.$
+If $F = \emptyset$, then the formulas have the form $\forall_XA\colon\emptyset \equiv \forall_XA \rightarrow false$ and $\exists_XA\colon\emptyset \equiv \exists_XA \& true$, since the empty disjunction is identical to $false$, whereas the empty conjunction is identical to $true$.
+The form $\forall_XA$ and $\exists_XA$ are abbreviations of the [abovementioned] formulas.
+If $X = \emptyset$, then $\forall A\colon F$ and $\exists A\colon F$ are analogous abbreviations.
 
-%language
-\subsection{The Language of Positively Constructed Formulas}
+The set of atoms $A$ is called {\em conjunct}. The empty conjunct is identical to $true$ as it was already mentioned.
 
-Let us denote the set of all conjuncts as $Con$ and assume that a {\em conjunct} is either a finite set of usual atoms of a first-order language, or $\boldsymbol{False}$, where $\boldsymbol{False}$ satisfies the property $A \subset \boldsymbol{False} $ for any $A \in Con$.
-The empty conjunct is denoted by $\boldsymbol{True}$.
-It is clear that if $A \in Con$ then $A \cup \boldsymbol{False} = \boldsymbol{False}$.
-Atoms of any conjunct (except $\boldsymbol{True}$ and $\boldsymbol{False}$) may contain variables, constants and function symbols.
+Variables from $X$ are bounded by correspoding quantifiers and called $\forall$--variables and $\exists$--variables, respectively.
+In $\forall_XA$ a variable from $X$ that does not appear in conjunct $A$ is called {\em unconfined} variable.
 
-\begin{definition}\label{def:1}
-Let $\bar{x}$ be a set of variables (possible empty set) and $A$ be a conjunct. Well-formed formulas of the PCF language are defined as follows:
+%В связи с изложенными сокращениями отметим следующий факт: 
+$\forall \emptyset \equiv \forall \emptyset\colon\emptyset \equiv \forall true \rightarrow false \equiv false$
 
-An expression $\exists \bar{x}\colon A$ is a $\exists$-formula; an expression $\forall \bar{x}\colon A$ is a $\forall$-formula.
+Construction $\forall_XA$ and $\exists_XA$ are called positive typical quantifiers (TQ), because $A$ is a conjunction of only positive atoms meaned as typical condition for $X$.
+In practise, this constructions denote the following phrases: ``for all $X$ satisfying $A$ there is...'' or ``there exists $X$ satisfying property $A$ such that...''.
+For example, ``for all integer $x,y,z$ and $n>2$ there is $x^n + y^n \ne z^n$''.
 
-Let $G_1,\ldots,G_k$ be a $\exists$-formulas, then $\forall$-formula is a formula of the following form: $\forall \bar{x}\colon A\left(G_1,\ldots,G_k\right)$.
+Originally, the term ``typical quantifier'' was introduced by N.
+Bourbaki \cite{Bourbaki} as part of notation for formalisation of mathematics.
+But typical quantifiers are stable in the languages of another applied fields.
 
-Let $G_1,\ldots,G_k$ be a $\forall$-formulas, then $\exists$-formula is a formula of the following form: $\exists \bar{x}\colon A\left(G_1,\ldots,G_k\right)$.
+%========================================================
+\section{The Method of PCF}
 
+\subsection{PCF Language}
+
+\begin{definition}[Positively--constructed formulas (PCF)]
+\label{def:pcf}
+Let, $X$ be the set of variables, and $A$ be a conjunct.
+\begin{enumerate}
+
+\item $\exists_XA$ and $\forall_XA$ are $\exists$--PCF and $\forall$--PCF respectively.
+
+\item If $F = \{F_1,\ldots,F_n\}$ are $\forall$--PCF, then $\exists_XA\colon F$ is a $\exists$--PCF.
+
+\item If $F = \{F_1,\ldots,F_n\}$ are $\exists$--PCF, then $\forall_XA\colon F$ is a $\forall$--PCF.
+
+\item Any $\exists$--PCF or $\forall$--PCF is a PCF.  
+\end{enumerate}
 \end{definition}
 
-An expression is a well--formed PCF if it is constructed based on above Definition \ref{def:1} rules only.
+This form of logical formulas is referred to as positively constructed formulas (PCFs), as they are written with only positive type quantifiers.
+The formulas contain no explicit logic negation sign.
+Any FOF can be represented as PCF \cite{ICDS2000}.
+%Таким образом ПО--формула есть особый вид записи классических формул языка предикатов, подобно КНФ, ДНФ и др.
 
-The variables from $\bar{x}$ are bound by a corresponding quantifiers, and these variables are called $\forall$-variables and $\exists$--variables correspondingly.
+PCF that started from $\forall \emptyset$ is called PCF in {\em canonical form}. Any PCF can be represented in the canonical form. Let $F$ is a $\exists$--PCF, then formula 
+$\forall \emptyset\colon F \equiv true \rightarrow F \equiv F$.
+If $F$ is a non-canonical $\forall$--PCF, then $\forall \emptyset\colon\{\exists \emptyset\colon F\} \equiv true \rightarrow \{true\&F\} \equiv F$.
+Typical quantifiers $\forall \emptyset$ and $\exists \emptyset$ are called {\em fictitious}, since they does not influence on truth of original PCF and does not bound any variables.
+%а только лишь служат конструкциями сохраняющими корректную запись ПОФ.
 
-%Any $\forall$--variable that does not occur in the corresponding conjunct is called {\em unconfined} variable.
+%Для удобства читаемости формул, будем представлять их в древовидной форме следующим образом:
+The PCFs are usually represented as a tree for more ease reading:
+$$Q_XA\colon\{F_1,\ldots,F_n\} \equiv Q_XA \left\{
+\begin{array}{lcl}
+ F_1 \\
+ \ldots \\
+ F_n
+\end{array}
+\right.,$$
 
-%semantic
-Let us define a semantic of PCFs by a common semantics of a corresponding first-order formula (FOF) in the classical first--order logic (FOL).
+\noindent{} where $Q$ is a quantifier.
+The tree elements have conventional names: node, root, leaf, branch etc.
+As the quantifiers $\forall$ correspond to disjunctions of formulas $\{F_1,\ldots,F_n\}$, (quantifiers $\exists$ correspond to a conjunction), then, in a general case, all $\forall$-nodes correspond a {\em disjunctive branching}, and for $\exists$-nodes correspond to {\em conjunctive branching}.
 
-\begin{definition}\label{def:semantic}
-Let $A = \{A_1,\ldots,A_l\}$ be a conjunct and $\bar{x} = \{x_1,\ldots,x_n\}$ be a set of variables.
-By $A^{\&}$ we denote $A_1 \&\ldots\&A_l$ and $\boldsymbol{False}^{\&}= False, \boldsymbol{True}^{\&}=True$ (propositional constants).
-By $F^{\text{\tiny{FOF}}}$ we denote an image of PCF formula $F$ in the FOL.
-
-%возможно убрать
-If $F= \exists \bar{x}\colon A$ then $F^{\text{\tiny{FOF}}} = \exists x_1 \ldots \exists x_n \left(A^{\&}\right).$
-%возможно убрать
-
-If $F = \forall \bar{x}\colon A$ then $F^{\text{\tiny{FOF}}} = \forall x_1 \ldots \forall x_n \left(A^{\&}\right).$
-
-If $F = \exists \bar{x}\colon A\left(G_1,\ldots,G_k\right)$ then $$F^{\text{\tiny{FOF}}} = \exists x_1 \ldots \exists x_n  \left(A^{\&} \& \left(G_{1}^{\text{\tiny{FOF}}} \& \ldots \& G_{k}^{\text{\tiny{FOF}}}\right)\right).$$
-
-If $F = \forall \bar{x}\colon A\left(G_1,\ldots,G_k\right)$ then $$F^{\text{\tiny{FOF}}} = \forall x_1 \ldots \forall x_n \left(A^{\&}\rightarrow \left(G_{1}^{\text{\tiny{FOF}}} \vee\ldots\vee G_{k}^{\text{\tiny{FOF}}}\right)\right).$$
-
-\end{definition}
-
-Any PCF can be obviously represented by a tree structure.
-Each node of that tree has the following form $\exists \bar{x}\colon A$ or $\forall \bar{x}\colon A$.
-Therefore, for readability we will represent PCF as a tree, as well as we will use corresponding terminology: node, branch, root, leaf, etc.
-%Каждый узел дерева имеет вид или $\forall\bar{x}\colon A$ or $\exists\bar{x}\colon A$.
-Например, в формуле вида $\forall \bar{x}\colon A\left(G_1,\ldots,G_k\right)$ корневым узлом являюется узел $\forall \bar{x}\colon A$, а его дочерние узлы это $G_1,\ldots,G_k\right$.
-
-If a PCF $F$ roots from the $\forall\colon\boldsymbol{True}$ node and each leaf of $F$ is a $\exists$-node then $F$ is called a PCF in the {\em canonical form}.
-It is clear, that any PCF $F$ can be converted to the canonical form by the following transformations:
+Some parts of canonical PCF are named as follows:
 \begin{enumerate}
-\item If $F$ is not canonical $\forall$-formula then $\forall\colon \boldsymbol{True}\left(\exists\colon \boldsymbol{True}\left(F\right)\right)$ is PCF rooted from the $\forall\colon\boldsymbol{True}$.
-\item If $F$ is $\exists$-formula then $\forall\colon \boldsymbol{True}\left(F\right)$ is PCF rooted from the $\forall\colon\boldsymbol{True}$.
-\item If $F$ has leaf $\forall \bar{x}\colon A$ then the node $\exists\colon\boldsymbol{False}$ can be added as child of this leaf.
+\item PCF root $\forall \emptyset$ is called {\em PCF root};
+\item PCF root children $\exists_XA$ are called {\em PCF bases}, conjunct $A$ is called {\em base of facts}, and PCF rooted from base is called {\em base subformula};
+\item PCF bases children $\forall_YB$ are caled {\em questions} to parent base. If question is a leaf of tree then it is called {\em goal question}. 
+\item subtrees of questions are called {\em consequents}. The consequent of goal question is $false$.
 \end{enumerate}
@@
 In the sequel, only PCFs in the canonical form are considered.
+%----------------------
+%EXAMPLE
+%----------------------
 
@@
 
-%--------------calculus
-\subsection{The Calculus of Positively Constructed Formulas}
+\subsection{Inference rule}
 
-The axiom schema $Ax\boldsymbol{JF}$ of the PCF calculus $\boldsymbol{JF}$ has the following form:
-$$\forall\colon\boldsymbol{True}\left(\exists \bar{x}_1\colon\boldsymbol{False}\left(\widetilde{\Phi}_1\right),\ldots,\exists \bar{x}_n\colon\boldsymbol{False}\left(\widetilde{\Phi}_n\right)\right).
-$$
-
-In the PCF calculus to prove $F$ we attempt to refute its negation, that's why the axioms are identical to false. Thus, an inference process in the PCF calculus is a {\em refutation} process.
-
-\begin{definition}
-\label{ircond}
-Let us assume that a question $\forall \bar{y}\colon A$ to a base $\exists \bar{x}\colon B$ has an answer $\theta$  if and only if $\theta$ is a substitution $\bar{y} \rightarrow H^{\infty}$ and $A\theta \subseteq B$, where $H^{\infty}$ is Herbrand universum based on variables from $\bar{x}$, constants and function symbols that occur in corresponding base subformula.
+\begin{definition}[Answer]
+\label{answer}
+Let us assume that a question $\forall_YB:\Upsilon$ to a base $\exists_XA$ has an {\em answer} $\theta$ if and only if $\theta$ is a substitution $Y \rightarrow H^{\infty} \cup X$ and $B\theta \subseteq A$, where $H^{\infty}$ is Herbrand universe based on constans and function symbols that occur in corresponding base subformula.
 \end{definition}
 
-%The unique unary inference rule $\bomega$ is defined as follows.
-
-\begin{definition}
-If a PCF $F$ has the structure $\forall\colon\boldsymbol{True}\left(\exists \bar{x}\colon B\left(\Phi\right),\Sigma\right)$, where $\Sigma$ is a list of other base subformulas, and $\Phi$ is a list of question subformulas, that contains a question subformula $\forall \bar{y}\colon A(\exists \bar{z}_i\colon C_i\left(\Psi_i\right))_{i=\overline{1,k}}$, then a formula $\bomega F$ is an application result of the unary inference rule $\bomega$ to the question $\forall \bar{y}\colon A$ with an answer $\theta$ and $\bomega F = \forall\colon\boldsymbol{True}(\exists \bar{x} \cup \bar{z}_i\colon B \cup C_i\theta\left(\Phi \cup \Psi_i\theta\right)_{i=\overline{1,k}},\Sigma).$
+\begin{definition}[Splitting]
+\label{splitting}
+Let $F = \exists_XA\colon\Psi$, and $S = \forall_YB\colon\Upsilon$, where $\Upsilon = \{\exists_{Z_1}C_1\colon\Gamma_1,\ldots,\exists_{Z_n}C_n\colon\Gamma_n\}$ then $split(F,S) = \{\exists_{X \cup {Z_1}'} A \cup {C_1}'\colon\Psi \cup {\Gamma_1}',\ldots,\exists_{X \cup {Z_n}'} A \cup {C_n}'\colon\Psi \cup {\Gamma_n}'\}$, where $'$ is a variable renaming operator.
+We say that $F$ splitted by $S$.
+Obviously, $split(F,\forall_YB) = split(F,\forall_YB\colon\emptyset) = \emptyset$.
 \end{definition}
 
-After appropriate renaming some bound variables inside each subformula the expression $\bomega F$ will satisfy the well--formed PCF requirements. %in the canonical form requirements
+\begin{definition}[$\omega$ inference rule]
+Consider some canonical PCF $F = \forall\emptyset\colon\Phi$.
+Let there exists a question $Q$ that has an answer $\theta$ to appropriate base $S \in \Phi$, then $\omega F  = \forall \emptyset\colon\Phi \setminus \{S\} \cup split(S,Q\theta)$.
+\end{definition}
 
-%
-Any finite sequence of PCFs $F, \bomega F, \bomega^2 F,\ldots,\bomega^n F$, where $\bomega^n F \in Ax\boldsymbol{JF}$, is called a {\em refutation} of $F$ in the PCF calculus.
+In other words, if a question has an answer to its base, then the base subformula is splitting by this question.
+In the case of the goal question, we say that the basic subformula is refuted because $split(S,\forall_YB) = \emptyset$.
+The refuted base subformula $S$ removed from the set of base subformulas $\Phi$, since $\Phi \setminus \{S\} \cup \emptyset = \Phi \setminus \{S\}$
 
-A question with $\exists:\boldsymbol{False}$ consequent node is called {\em goal question}. If a question has disjunctive branching then answering this question results in formula {\em splitting}.
+Once refuted all bases subformulas from $\Phi$, the formula $F$ is also refuted, since $\forall \emptyset\colon\emptyset \equiv false$. %согласно правилу описанному выше
 
-For clarity consider simple example of a PCF refutation.
+%Note that the answer the questin with the disjunctive branching leads to an increase
 
+%Отметим, что ответ на вопрос с дизъюнктивным ветвлением приводит к увеличению количества базовых подформул в силу использования оператора расщепления, а опровержение базы уменьшает количество баз на одну опровергнутую.
+
+%----------------------
+%EXAMPLE
+%----------------------
 \begin{example}[A refutation in $\boldsymbol{JF}$]\label{proofexample}

### 2015-08-24 paper.tex
Title slightly corrected.

@@
 % first the title is needed
-\title{Automated Theorem Proving Based on the Calculus of Positively Constructed Formulas}
+\title{Automated Theorem Proving Based on the Calculus of Positively Constructed Formulas\thanks{Ok}}
 % a short form should be given in case it is too long for the running head
@@
 %
-\author{Alexander Larionov%
-\thanks{Ok}%
-\and Artem Davydov\and Evgeny Cherkashin\and Ivan Terekhin}
+\author{Alexander Larionov\and Artem Davydov\and Evgeny Cherkashin}
 %
@@
               1, Karl Marks str., Irkutsk, Russia \\
-           Institute for System Dynamics and Control Theory,
-              134, Lermontov str., Irkutsk, Russia   
+           Institute for System Dynamics and Control Theory at Siberian Branch
+              of Russian Academy of Sciences,
+              134, Lermontov str., Irkutsk, Russia\\
+           Irkutsk National Research Technical University,
+              83, Lermontov str., Irkutsk, Russia
 %\mailsa\\
@@
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2015-08-24 paper.tex
Reviewed the language definition.

@@
 % first the title is needed
-\title{Automated Theorem Proving Based on the Calculus of Positively Constructed Formulas\thanks{Ok}}
+\title{Software for Automated Theorem Proving Based on the Calculus of Positively Constructed Formulas\thanks{Ok}}
 % a short form should be given in case it is too long for the running head
@@
 \begin{abstract}
-A description of the calculus of positively constructed formulas (PCF) and prover based on this calculus is considered in this paper.
+A description of the calculus of positively constructed formulas (PCF) and prover based on this calculus is considered.
 The PCF calculus has been developed by Russian scientists S.N.
@@
 The prover is being tested by means of solving problems from TPTP library.
-The usage of implemented strategies is also commented in this paper.
+The usage of implemented inference search strategies is also commented in this paper.
 \keywords{positively constructed formulas, automated theorem proving, proof strategies}
@@
 Zherlov by an evolutionary way in describing and solving control theory (CT) problems.
-In \cite{ICDS2000} the PCF calculus is presented as first-order logical formalism, examples of CT problems described  and solved by the PCF calculus (elevator group control, mobile robot action planning and telescope guidance), proof of soundness and completeness are presented as well.
+In \cite{ICDS2000} the PCF calculus is presented as first-order logical formalism, examples of CT problems described and solved by the PCF calculus (elevator group control, mobile robot action planning and telescope guidance), as well as proof of soundness and completeness.
 
-The PCF calculus is positioned as not only machine-oriented, but also as a human-oriented, mostly aimed at solving the problems of dynamic systems control due to the following features of this calculus: unique inference rule and simple scheme of axioms; modifiability of semantics (constructive, monotonic, temporal, etc.) and besides it is possible to construct intuitionistic inferences of some non-Horn formulas; the explicit usage of $\forall$ and $\exists$ quantifiers, the scolemization procedure is not required.
+The PCF calculus is both machine-oriented, and also human-oriented, naturally aimed at solving the problems of dynamic systems control thanks to its features such as follows: unique inference rule and simple scheme of axioms; modifiability of semantics (constructive, monotonic, temporal, etc.) and besides it is possible to construct intuitionistic inferences of some non-Horn formulas; the explicit usage of $\forall$ and $\exists$ quantifiers, the scolemization procedure is not required.
 
 However, in this paper, we will not cover the human--orientability properties of the PCF calculus and its dynamic systems control properties as well.
-We deal with the automatic search of a logical inference, as the inference engine capabilities are the basis of the problem solving systems, described in \cite{ICDS2000}.
+This and the calculus capabilities in action planning is described in \cite{ICDS2000}.
+Here, we deal with the automatic search of a logical inference, i.e., the inference engine capabilities.
 In order to estimate the applicability of the PCF calculus for automatic theorem proving we develop a prover program and test it on problems from TPTP library.
@@
 %=======================================================================
-\section{Background}
-We consider the language of first-order logic consist in formulas build on atomic formulas with $\&, \vee, \neg, \rightarrow, \leftrightarrow$ operators, $\forall$ and $\exists$ quantifier symbols and constant operators $true$ and $false$.
-The terms term, atom, literal we define in the usual way.
-FOFs.
+\section{Preliminaries PCF Calculus}
 
-Let $X = \{x_1,\ldots,x_k\}$ be the set of variables, $A = \{A_1,\ldots,A_m\}$ be the set of atomic formulas, and $F = \{F_1,\ldots,F_n\}$ be the set of some FOFs.
-Then the following formulas $((\forall x_1) \ldots (\forall x_k) (A_1 \& \ldots \& A_m \rightarrow (F_1 \vee \ldots \vee F_n)))$ and $((\exists x_1) \ldots (\exists x_k) (A_1 \& \ldots \& A_m \& (F_1 \& \ldots \& F_n)))$ are denoted as  $\forall_XA\colon F$ and $\exists_XA\colon F$ respectively, keeping in mind that the $\forall$--quantifier corresponds to $\rightarrow F$, where $F$ means disjunction of all formulas from $F$, and $\exists$--quantifier corresponds to $\& F$, where $F$ means conjunctions of all formulas from $F$.
+Let's consider a language of first-order logic (FOL) that consists of first-order formulas (FOFs) built out of atomic formulas with $\&, \vee, \neg, \rightarrow, \leftrightarrow$ operators, $\forall$ and $\exists$ quantifier symbols and constants $true$ and $false$.
+The concepts of \emph{term}, \emph{atom}, \emph{literal} we define in the usual way.
+
+Let $X = \{x_1,\ldots,x_k\}$ be a set of variables, $A = \{A_1,\ldots,A_m\}$ be a set of atomic formulas, and $F = \{F_1,\ldots,F_n\}$ be a set of FOFs.
+Then the following formulas $((\forall x_1) \ldots (\forall x_k) (A_1 \& \ldots \& A_m \rightarrow (F_1 \vee \ldots \vee F_n)))$ and $((\exists x_1) \ldots (\exists x_k) (A_1 \& \ldots \& A_m \& (F_1 \& \ldots \& F_n)))$ are denoted as  $\forall_XA\colon F$ and $\exists_XA\colon F$ respectively, keeping in mind that the $\forall$--quantifier corresponds to $\rightarrow F^{\vee}$, where $F^{\vee}$ means disjunction of all formulas from $F$, and $\exists$--quantifier corresponds to $\& F^{\&}$, where $F^{\&}$ means conjunctions of all formulas from $F$.
 
 If $F = \emptyset$, then the formulas have the form $\forall_XA\colon\emptyset \equiv \forall_XA \rightarrow false$ and $\exists_XA\colon\emptyset \equiv \exists_XA \& true$, since the empty disjunction is identical to $false$, whereas the empty conjunction is identical to $true$.
-The form $\forall_XA$ and $\exists_XA$ are abbreviations of the [abovementioned] formulas.
+The form $\forall_XA$ and $\exists_XA$ are abbreviations of such formulas.
 If $X = \emptyset$, then $\forall A\colon F$ and $\exists A\colon F$ are analogous abbreviations.
@@
 
-Variables from $X$ are bounded by correspoding quantifiers and called $\forall$--variables and $\exists$--variables, respectively.
-In $\forall_XA$ a variable from $X$ that does not appear in conjunct $A$ is called {\em unconfined} variable.
+Variables from $X$ are bound by correspoding quantifiers and called $\forall$--variables and $\exists$--variables, respectively.
+In $\forall_XA$, a variable from $X$ that does not appear in conjunct $A$ is called {\em unconfined} variable.
 
@@
 
-Construction $\forall_XA$ and $\exists_XA$ are called positive typical quantifiers (TQ), because $A$ is a conjunction of only positive atoms meaned as typical condition for $X$.
-In practise, this constructions denote the following phrases: ``for all $X$ satisfying $A$ there is...'' or ``there exists $X$ satisfying property $A$ such that...''.
-For example, ``for all integer $x,y,z$ and $n>2$ there is $x^n + y^n \ne z^n$''.
+Construction $\forall_XA$ and $\exists_XA$ are called positive \emph{type quantifiers} (TQ), because $A$ is a conjunction of only positive atoms and referred to as also as \emph{type condition} for $X$.
+In practise, this constructions denote phrases such as follows: ``for all $X$ satisfying $A$ there is...'' or ``there exists $X$ satisfying property $A$ such that...''; ``for all integer $x,y,z$ and $n>2$ there is $x^n + y^n \ne z^n$''.
 
-Originally, the term ``typical quantifier'' was introduced by N.
-Bourbaki \cite{Bourbaki} as part of notation for formalisation of mathematics.
-But typical quantifiers are stable in the languages of another applied fields.
+Originally, the term ``type quantifier'' was introduced by N.~Bourbaki \cite{Bourbaki} as part of notation for formalization of mathematics.
+But type quantifiers are stable in languages of another applied fields.
 
@@
 \label{def:pcf}
-Let, $X$ be the set of variables, and $A$ be a conjunct.
+Let, $X$ be a set of variables, and $A$ be a conjunct.
 \begin{enumerate}
@@
 
-PCF that started from $\forall \emptyset$ is called PCF in {\em canonical form}. Any PCF can be represented in the canonical form. Let $F$ is a $\exists$--PCF, then formula
+PCF that started from $\forall \emptyset$ is called PCF in a {\em canonical form}. Any PCF can be represented in the canonical form. Let $F$ is a $\exists$--PCF, then formula
 $\forall \emptyset\colon F \equiv true \rightarrow F \equiv F$.
 If $F$ is a non-canonical $\forall$--PCF, then $\forall \emptyset\colon\{\exists \emptyset\colon F\} \equiv true \rightarrow \{true\&F\} \equiv F$.
-Typical quantifiers $\forall \emptyset$ and $\exists \emptyset$ are called {\em fictitious}, since they does not influence on truth of original PCF and does not bound any variables.
+Type quantifiers $\forall \emptyset$ and $\exists \emptyset$ are called {\em fictitious}, since they do not influence on truth value of an original PCF and do not bind any variables.
 %а только лишь служат конструкциями сохраняющими корректную запись ПОФ.
@@
 %Для удобства читаемости формул, будем представлять их в древовидной форме следующим образом:
-The PCFs are usually represented as a tree for more ease reading:
+The PCFs are usually represented as trees for more ease reading:
 $$Q_XA\colon\{F_1,\ldots,F_n\} \equiv Q_XA \left\{
@@
 
-\noindent{} where $Q$ is a quantifier.
-The tree elements have conventional names: node, root, leaf, branch etc.
-As the quantifiers $\forall$ correspond to disjunctions of formulas $\{F_1,\ldots,F_n\}$, (quantifiers $\exists$ correspond to a conjunction), then, in a general case, all $\forall$-nodes correspond a {\em disjunctive branching}, and for $\exists$-nodes correspond to {\em conjunctive branching}.
+\noindent where $Q$ is a quantifier.
+Tree elements have conventional names: \emph{node}, \emph{root}, \emph{leaf}, \emph{branch}, etc.
+As the quantifiers $\forall$ correspond to disjunctions of formulas $\{F_1,\ldots,F_n\}$ (quantifiers $\exists$ correspond to a conjunction), then all $\forall$-nodes correspond a {\em disjunctive branching}, and for $\exists$-nodes correspond to {\em conjunctive branching}.
 
-Some parts of canonical PCF are named as follows:
+Some parts of canonical PCF are denoted as follows:
 \begin{enumerate}
-\item PCF root $\forall \emptyset$ is called {\em PCF root};
-\item PCF root children $\exists_XA$ are called {\em PCF bases}, conjunct $A$ is called {\em base of facts}, and PCF rooted from base is called {\em base subformula};
-\item PCF bases children $\forall_YB$ are caled {\em questions} to parent base. If question is a leaf of tree then it is called {\em goal question}.
-\item subtrees of questions are called {\em consequents}. The consequent of goal question is $false$.
+\item PCF root $\forall \emptyset$ is called PCF {\em root};
+\item Each PCF root child $\exists_XA$ is called PCF {\em base}, conjunct $A$ is called {\em base of facts}, and PCF rooted from base is called {\em base subformula};
+\item PCF base children $\forall_YB$ are called {\em questions} to its parent base.  If a question is a leaf of a tree then it is called {\em goal question}.
+\item subtrees of questions are called {\em consequents}.  The absent consequent of a goal question is identical to $false$.
 \end{enumerate}
@@
 
-Some parts of a PCF are denoted by particular terms: the topmost (0-depth) node is called {\em root}; a 1-depth node is called {\em base} (of formula); a maximal subtree that rooted at the 1-depth node is called {\em base subformula}; a 2-depth node is called {\em question} (to its base); a maximal subtree that rooted at the 2-depth node is called {\em question subformula}; a maximal subtree that rooted at 3-depth node is called {\em consequent} (subformula).
+% Some parts of a PCF are denoted by particular terms: the topmost (0-depth) node is called {\em root}; a 1-depth node is called {\em base} (of formula); a maximal subtree that rooted at the 1-depth node is called {\em base subformula}; a 2-depth node is called {\em question} (to its base); a maximal subtree that rooted at the 2-depth node is called {\em question subformula}; a maximal subtree that rooted at 3-depth node is called {\em consequent} (subformula).
 With respect to Definition~\ref{def:semantic} if an even-depth (odd-depth) node has more than one children then the node has {\em disjunctive branching} (correspondingly {\em conjunctive branching}).%; if a node of PCF has a following form $\forall \bar{x}\colon A$ and if $\exists v \in \bar{x}$ and $v \notin A$ then variable $v$ is called {\em unconfined} variable.
@@
 \begin{example}
-Let us consider a formula of FOL.
+Let us consider a PCF representation of a FOF.
 $$F= \neg\bigl(\forall x\:\exists y P(x,y)\rightarrow \exists z P(z,z)\bigr).$$

### 2015-09-04 paper.tex
Changed parentsis to curly brackets, removed redundants.
Corrected constraint strategies.

@@
 
-$$F^{\text{\tiny{PCF}}} = \forall\colon \boldsymbol{True}(\exists\colon\boldsymbol{True}(\forall x\colon\boldsymbol{True}(\exists y\colon P(x,y)), \forall z\colon P(z,z)(\exists\colon\boldsymbol{False}))).$$
+$$F^{\text{\tiny{PCF}}} = \forall\colon \boldsymbol{True}\{\exists\colon\boldsymbol{True}\{\forall x\colon\boldsymbol{True}\{\exists y\colon P(x,y)\}, \forall z\colon P(z,z)\{\exists\colon\boldsymbol{False}\}\}\}.$$
 %А эта ПО-формула в древовидной форме имеет следующий вид+++++++++++++++++++++++++
@@
 %Для удобствоа мы используем символ --- для обозначения рёбер в древовидном представлении ПО-формулы. Фугрная скобка также является вспомогательным символом.+++++++++++++++++++++++++
-For convenience we use symbols ``--'' and ``\{'' for denoting the edges and branching in tree--like representation of PCF.
+For convenience we use symbol ``--'' to denote the adjacent edges in tree--like representation of PCF.
 
-\subsection{Inference rule}
+\subsection{The Inference rule}
 
@@
 \label{answer}
-Let us assume that a question $\forall_YB:\Upsilon$ to a base $\exists_XA$ has an {\em answer} $\theta$ if and only if $\theta$ is a substitution $Y \rightarrow H^{\infty} \cup X$ and $B\theta \subseteq A$, where $H^{\infty}$ is Herbrand universe based on constans and function symbols that occur in corresponding base subformula.
+A question $\forall_YD:\Upsilon$ to a base $\exists_XA$ has an {\em answer} $\theta$ if and only if $\theta$ is a substitution $Y \rightarrow H^{\infty} \cup X$ and $D\theta \subseteq A$, where $H^{\infty}$ is Herbrand universe based on constans and function symbols that occur in corresponding base subformula.
 \end{definition}
@@
 \label{splitting}
-Let $F = \exists_XA\colon\Psi$, and $S = \forall_YB\colon\Upsilon$, where $\Upsilon = \{\exists_{Z_1}C_1\colon\Gamma_1,\ldots,\exists_{Z_n}C_n\colon\Gamma_n\}$ then $split(F,S) = \{\exists_{X \cup {Z_1}'} A \cup {C_1}'\colon\Psi \cup {\Gamma_1}',\ldots,\exists_{X \cup {Z_n}'} A \cup {C_n}'\colon\Psi \cup {\Gamma_n}'\}$, where $'$ is a variable renaming operator.
-We say that $F$ splitted by $S$.
-Obviously, $split(F,\forall_YB) = split(F,\forall_YB\colon\emptyset) = \emptyset$.
+Let $B = \exists_XA\colon\Psi$, and $Q = \forall_YD\colon\Upsilon$, where $\Upsilon = \{\exists_{Z_1}C_1\colon\Gamma_1,\ldots,\exists_{Z_n}C_n\colon\Gamma_n\}$ then $split(B,Q) = \{\exists_{X \cup {Z_1}'} A \cup {C_1}'\colon\Psi \cup {\Gamma_1}',\ldots,\exists_{X \cup {Z_n}'} A \cup {C_n}'\colon\Psi \cup {\Gamma_n}'\}$, where $'$ is a variable renaming operator.
+We say that $B$ is split by $Q$.
+Obviously, $split(B,\forall_YD) = split(B,\forall_YD\colon\emptyset) = \emptyset$.
 \end{definition}
 
-\begin{definition}[$\omega$ inference rule]
+\begin{definition}[the inference rule $\bomega$]
 Consider some canonical PCF $F = \forall\emptyset\colon\Phi$.
-Let there exists a question $Q$ that has an answer $\theta$ to appropriate base $S \in \Phi$, then $\omega F  = \forall \emptyset\colon\Phi \setminus \{S\} \cup split(S,Q\theta)$.
+Let there exists a question $Q$ that has an answer $\theta$ to appropriate base $B \in \Phi$, then $\omega F  = \forall \emptyset\colon\Phi \setminus \{B\} \cup split(B,Q\theta)$.
 \end{definition}
 
-In other words, if a question has an answer to its base, then the base subformula is splitting by this question.
-In the case of the goal question, we say that the basic subformula is refuted because $split(S,\forall_YB) = \emptyset$.
-The refuted base subformula $S$ removed from the set of base subformulas $\Phi$, since $\Phi \setminus \{S\} \cup \emptyset = \Phi \setminus \{S\}$
+In other words, if a question has an answer to its base, then the base subformula is split by this question.
+In the case of a goal question, we say that the basic subformula is refuted because $split(B,\forall_YD) = \emptyset$.
+The refuted base subformula $B$ removed from the set of base subformulas $\Phi$, since $\Phi \setminus \{S\} \cup \emptyset = \Phi \setminus \{S\}$
 
-Once refuted all bases subformulas from $\Phi$, the formula $F$ is also refuted, since $\forall \emptyset\colon\emptyset \equiv false$. %согласно правилу описанному выше
+As soon as all the bases subformulas from $\Phi$ have been refuted, the formula $F$ is also refuted, since $\forall \emptyset\colon\emptyset \equiv false$.
+The PCF language and the inference rule $\bomega$ form the calculus $\boldsymbol{JF}$ oriented to refutation of a negation of an original formula.
+The only $\boldsymbol{JF}$--axiom is $\forall \emptyset\colon\emptyset$, i.e., \emph{false}.
 
-%Note that the answer the questin with the disjunctive branching leads to an increase
+%согласно правилу описанному выше
+
+%Note that the answer the question with the disjunctive branching leads to an increase
 
@@
 \begin{equation*}\label{ex:f1}
-  F_1 = \forall\colon\boldsymbol{True} - \exists\colon S(e)(Q_1,Q_2,Q_3,Q_4);
+  F_1 = \forall\colon\boldsymbol{True} - \exists\colon S(e)\{Q_1,Q_2,Q_3,Q_4\};
 \end{equation*}
@@
 \end{equation*}
-At the first step only one answer $\{x \rightarrow e\}$ to question $Q_1$ exists. After applying rule $\bomega$ with this answer, formula will be in the following form:
+At the first step there is only one answer $\{x \rightarrow e\}$ to question $Q_1$.
+After applying rule $\bomega$ with this answer to the only base $F_1$, the formula will be converted to the following form:
 \begin{equation*}\label{ex:f2}
-  F_2 = \forall\colon\boldsymbol{True} - \exists\colon S(e),A(a)(Q_1,Q_2,Q_3,Q_4).
+  F_2 = \forall\colon\boldsymbol{True} - \exists\colon S(e),A(a)\{Q_1,Q_2,Q_3,Q_4\}.
 \end{equation*}
-At the second step also only one answer $\{x \rightarrow a\}$ to  question $Q_4$ exists.
+At the second step there is also only one answer $\{x \rightarrow a\}$ to question $Q_4$.
 After applying $\bomega$ with this answer, formula is split, because $Q_4$ has disjunctive branching.
-And now formula has the following form:
+The formula will have the following form:
 
@@
 
-At the third step first base can be refuted by answering on $Q_3$ goal question with $\{x \rightarrow y_1; y \rightarrow a\}$.
-Refuted base (base subformula) can be deleted from the list of base subformulas.
+At the third step the first base can be refuted by answering on $Q_3$ (goal question) with $\{x \rightarrow y_1; y \rightarrow a\}$.
+The refuted base (and its whole base subformula) is to be deleted from the list of base subformulas.
 
-At the fourth step the answer $\{z \rightarrow a\}$ to fifth new question exists. And now formula has the following form:
+At the fourth step there is the answer $\{z \rightarrow a\}$ to fifth new question.  The formula will be of the following form:
 \begin{equation*}\label{ex:f5}
@@
 
-At the fifth step the only base can be refuted by answering on $Q_4$ goal question with  answer $\{x \rightarrow a; y \rightarrow f(a)\}$.
+At the fifth step the only base can be refuted by answering on $Q_4$ (goal question) with  answer $\{x \rightarrow a; y \rightarrow f(a)\}$.
 
-The refutation is finished because all bases were refuted.
+The refutation is finished because all the bases were refuted.
 
@@
 Our prover is implemented in the D programming language \cite{DPL1,DPL2}.
-The D programming language has many features: compiler produce a native binary code, which both fast and optimized; procedures of memory management are manually adjustable, but it is also possible to use garbage collection; there is a good library for parallel computation construction in Erlang--like model, which is generally recognized as the quite good \cite{Erlang1}.
+The language has many convenient features: compiler produce a native binary code, which both fast executed and optimized by its compiler; memory management is manually adjustable, but it is also possible to use adjustable garbage collection techniques implemented in translator; there is a good library for parallel computation construction in the image and likeness of Erlang, which is generally recognized as the reasonably good \cite{Erlang1}.
 
@@
 
-Each step of the inference in the PCF calculus can be considered as an addition of corresponding consequent subformulas to the base subformula (facts to a base; questions to a question lists).
-Thus, a formula is usually monotonically increasing, and splitting if a disjunctive branching happens.
+Each step of the inference in the PCF calculus can be considered as addition of special cases of corresponding consequent subformulas to the base subformula (facts to a base; questions to a question sets).
+Thus, a formula size will monotonically increase due to splitting of disjunctive branching.
 %---------
 
-One of the main tools of the logical inference search implementation in the developed prover is the \emph{proof state tree} (PST), which is built with the data structures based on so--called chunks described below.
-The PST structures are used to store the sequence of the steps of logical inference.
+One of the main tools of the logical inference search implementation of the developed prover is the \emph{proof state tree} (PST) data structure.
+PST is constructed from so--called chunk data structures described below.
+The PST is used to store the sequence of the steps of a logical inference.
 Each step corresponds to a single operation that were performed.
 PST also represents the current state of the formula under refutation.
-The main purpose of PST is to strictly fix all operations and their parameters, which occurs on every step of logical inference construction.
-The example of such operation is the fact of applying a substitution in a base subformula to a question.
+The main purpose of PST is to strictly fix all performed operations and their parameters, which are accompanied every step of logical inference construction.
+An application of a substitution in a base subformula to a question is an example of such operation.
 
 %--------------------------------------------------------------------------------
-\subsubsection{Chunk.}
-Chunk is a single--linked list of elements with type {\tt T} which have not only reference to the \texttt{first} and \texttt{last} elements of that list.
+\subsubsection{The chunk data structure.}
+The \emph{chunk} is a single--linked list of elements of type {\tt T}.
+The \texttt{first} and the \texttt{last} elements of the list are distinguished.
 A chunk is empty, if it does not contain any elements, and \texttt{first} and \texttt{last} elements are NULL--references.
-We say, that chunk $C_1$ is \emph{linked} with chunk $C_2$ \emph {from the left}, if the \texttt{last} node of chunk $C_1$ is referring to the \texttt{first} node of chunk $C_2$ and, simultaneously, $C_2$ linked with $C_1$ \emph{from the right}.
-For this operation a procedure of linking {\tt void link(Chunk!(T) c)} is defined, which links current chunk and chunk {\tt c} from left.
-If the chunks are not empty, then linking is proceeded according to the definition.
+We say, that chunk $C_1$ is \emph{linked} to chunk $C_2$ \emph {from the left hand}, if the \texttt{last} node of chunk $C_1$ is referring to the \texttt{first} node of chunk $C_2$, the linking ``\emph{from the right hand}'' is similar.
+Linking operations are implemented in D--language procedure {\tt void link(Chunk!(T) c)}, which links current chunk and chunk {\tt c} from left.
+If the chunks are not empty, then linking is made according to the definition.
 If one of the chunks is empty, then its \texttt{first} and \texttt{last} elements will be assigned to the value of the linking element.
-Note, that several chunks can be linked with different chunk from one of its sides simultaneously.
-Therefore, using the procedure of linking one can build a tree of chunks, which grows from leafs.
+Note, that several chunks can be linked with the same chunk from the same side simultaneously.
+Therefore, one can build a tree of chunks growing from leafs.
 
@@
 
-In Fig.~\ref{fig:chunk1} an example of structure is presented, which is built from chunks.
+In Fig.~\ref{fig:chunk1} an example of a structure is presented, which is built from chunks.
 \begin{figure}[h]
@@
   \includegraphics[width=0.6\linewidth]{img/ChunkFullp1.eps}
-  \caption{Chunks}
+  \caption{An usage of chunk data structures}
   \label{fig:chunk1}
 \end{figure}
-There are five chunks in this figure, and several of them are linked together with \texttt{first} field: ``5'' and ``4'', ``4'' and ``3'', ``3'' and ``1'', ``2'' and ``1''.
+There are five chunks in this figure, and several of them are linked together with \texttt{first} element: ``5'' and ``4'', ``4'' and ``3'', ``3'' and ``1'', ``2'' and ``1''.
 Chunk 4 is empty.
-\texttt{first} and \texttt{last} is the first and last nodes of chunks.
 Numbers are the identifiers of the corresponding nodes.
-Empty chunk duplicates node \texttt{first} for the node linked with empty chunk from the left.
+Empty chunk duplicates node \texttt{first} of its left hand linked node.
 
@@
 \subsubsection{Proof state tree.}
-A proof state tree is a tree with the following properties: the root is the original PCF, other nodes are consequent subformulas (with the answer applied to them and variables renamed).
+A proof state tree (PST) is a tree structure of the following properties: the root is the original PCF, other nodes are consequent subformulas (with its answer applied and variables renamed).
 In addition, the nodes contain some system information like step of inference, answer, etc.
-Thus, the number of the leafs is the current number of the bases, and at current inference step a PCF is denoted by the path from the leaf to the PST root.
+Thus, the number of the leafs of the PST is the current number of the bases, and at current inference step, a PCF is denoted by the path from the leaf to the PST root.
 This approach allows us to rollback (backtrack) the inference search process, observe proof states and share references among terms and formulas.
-If a base is refuted then its memory can be released by deleting of the path from corresponding leaf to the soonest branching.
+If a base is refuted then its structures can be released by deleting the path from corresponding leaf to the nearest branching.
 
-Every node of PST is a chunk. The growth of the tree is a linking of new chunks with the existing leafs of the tree from the left.
+Every node of a PST is a chunk. The growth of the tree is a linking of new chunks with the existing leafs of the tree from the left hand.
 
@@
 For example, formulas that are denoted by paths ``5''--``1'' and ``3''--``1'' share data that are represented by nodes ``1''--``2''.
-When base subformula that is represented by path ``3''--``1'' is refuted, we can delete path from node ``3'' to soonest branching --- in this case only node ``3'' is deleted, because nodes ``2''--``1'' are still used for representing another base subformulas.
+When base subformula that is represented by path ``3''--``1'' is refuted, we can delete path from node ``3'' to the nearest branching.
+In this case only node ``3'' is deleted because nodes ``2''--``1'' are still used for representing another base subformulas.
 %The lettering in figures should have a height of 2mm (10-point type)
@@
   \includegraphics[width=0.4\linewidth]{img/PSTp1.eps}
-  \caption{Proof state tree for PCF from Example \ref{proofexample}.}
+  \caption{Proof state tree for PCF from Example \ref{proofexample}}
   \label{fig:pst}
@@
 %------
-In addition, in a PST the data--sharing (memory sharing) is implemented: if two base subformulas are characterized by the corresponding paths, their bases share data, which are represented by the same part of these paths.
+The PST structure allow to implement the memory data sharing: if base subformulas have common PST paths, their bases share elements of type conditions and sets of questions (Fig.~\ref{fig:datasharing1}).
 \begin{figure}[h]
@@
   \includegraphics[width=0.4\linewidth]{img/Graphsp1.eps}
-  \caption{Example of proof state tree for base subformula refutation}
+  \caption{Memory data sharing}
   \label{fig:datasharing1}
@@
 
-If a PCF contains a question with disjunctive branching, then its base subformula will be divided into several new base subformulas during the searching procedure of the logical inference.
+If a PCF contains a question with disjunctive branching, then its base subformula will be divided into several new base subformulas during searching procedure of the logical inference.
 The number of new subformulas corresponds to the number of disjunctive subformulas in the consequent of the question.
-In the straightforward variant of the prover implementation, the sharing will require copying previous state of the original formula more than once; this copying have a linear complexity, but it leads to the redundant expenses of memory and processor time.
-PST allows one to implement sharing the common subbranches.
-Let's we have the following base subformula:
+In a straightforward variant of the prover implementation, the sharing will require copying previous state of the original formula more than once; this copying have a linear complexity, and it leads to the redundant expenses of memory and processor time.
+PST allows us to implement common subbranches sharing instead of copying.
+Let's consider the following base subformula:
 $$\exists: A(a) - \forall x: A(x) - \left\{
@@
  \exists \colon B(x) & - & \forall y: B(y) - \exists\colon\boldsymbol{False}\\
- \exists \colon C(x) & - & \forall y: C(y) - \exists\colon\boldsymbol{False}
+ \exists \colon C(x) & - & \forall y: C(y) - \exists\colon\boldsymbol{False}.
 \end{array}
 \right. $$
-
-Note, that the formula has disjunctive branching and it is deep. PST for the refutation of this base subformula is presented on Fig.~\ref{fig:datasharing2}
+The formula has disjunctive branching and it is deep. PST for the refutation of this base subformula is presented in Fig.~\ref{fig:datasharing2}.
 \begin{figure}[h]
@@
   \includegraphics[width=0.82\linewidth]{img/DataSharing2p1.eps}
-  \caption{Example of proof state tree for base subformula refutation}
+  \caption{Example of PST for a base subformula refutation}
   \label{fig:datasharing2}
@@
 %====================================================================================
-\subsection{Lazy concretization}
+\subsection{Lazy instantiation}
 
-By the Definition~3 the inference rule $\omega$ can be applied if the question $\forall \bar{y}\colon A$ to base $\exists \bar{x}\colon B$ has an answer $\theta$, such that $\theta$ is a substitution $\bar{y} \rightarrow H^{\infty}$ and $A\theta \subseteq B$.
-If all the variables in the $\bar{y}$ are occurred in $A$, then the search of corresponding elements from $H^{\infty}$ is carried on by the algorithm of subsumption and depends on only corresponding base of facts, hence the conditions of sets subsumptions need to be satisfied.
-If a variable in the $\bar{y}$ is not occurred in $A$ then the variable is called {\emph unconfined variable} and it is not used in a subsumption.
-However, we need to find a substitution for these variables as well, and any element of Herbrand universum is a correct substitution.
-The problem with Herbrand universum is that in the general case, it is an unlimited set and it is not known what element should be taken as a substitution.
+According to the Definition~3, the inference rule $\omega$ can be applied if a question $\forall \bar{y}\colon A$ to a base $\exists \bar{x}\colon B$ has an answer $\theta$, such that $\theta$ is a substitution $\bar{y} \rightarrow H^{\infty}$ and $A\theta \subseteq B$.
+If all the variables in $\bar{y}$ are occurred in $A$, then the search of corresponding elements from $H^{\infty}$ is carried on by the algorithm of subsumption among corresponding base of facts, since the conditions of subsumption of the sets have to be satisfied.
+If a variable in the $\bar{y}$ is not occurred in $A$ then the variable is called \emph{unconfined variable} and it is not used in a subsumption.
+However, we need to find a substitution for these variables as well.
+Each of the elements of Herbrand universe is a correct substitution for the unconfined variable.
+This rises a problem related to Herbrand universe.
+In the general case the universe is an unlimited set, and it is not known what element should be taken as a substitution for the variable.
 
-The main concept of lazy concretization is as follows.
-Instead of selection of a concrete element of the Herbrand universum for a substitution of an unconfined variable an {\em unspecified Herbrand element} (UHE) is used temporally.
+The main concept of lazy instantiation is as follows.
+Instead of selecting a concrete element of a Herbrand universum for a substitution of an unconfined variable an {\em unspecified Herbrand element} (UHE) is used temporally.
 The UHE will be concretized further to a proper term.
-Such concretization is based on the logical inference requirements.
-UHE will be concretized to a ground term, or in some situations, it will remain unconcretized.
-In such case, the required terms provide the possibility of answering to the question: UHE is gradually concretized in such way, that on every step of inference we can build new answer for some question.
-Simultaneously the following problem is being solved: to what term we need to concretize UHE to make it possible to answer the question?
-Note, that originally UHE appears as a part of answer in one question and its concretization is done while searching the answers for other questions.
-As ``gradually concretization'' we mean procedure of lazy computations: UHE concretized as precise, as it is enough for answering the current question, i.e.
-concretization, generally, is partial, for example UHE for $h$ is concretized to $f(h_1)$, where $h_1$ is a new UHE.
+The result of the concertizing depends on the logical inference strategy.
+UHE will be instantiated to a ground term, or in some situations, it could remain unbound.
 
-By it's nature UHE is similar to unconcretized $\forall$-variable in the sense that it is mutable, i.e.
-it is potentially concretized, however all such changes must be directed only to concretization of UHE, i.e.
-UHE is substituted only to some term or to different UHE, but not to the variable.
+The UHE can be concretized in a term, containing other UHEs.
+The variant of the concretizing on the each step is chosen according to inference strategies to make refutation possible, e.g.
+by analyzing a term structure of a question.
+The process is organized in a trendy ``lazy'' manner: UHE is instantiate as concrete, as it is enough for answering a question, i.e.
+instantiation is generally partial, for example UHE for $h$ is concretized to $f(h_1)$, where $h_1$ is a new UHE.
 
-Let's consider the example of this technique in process of building the logical inference for the following formula.
+UHE by nature is similar to unconcretized $\forall$-variable in the sense that it is mutable, i.e., it is potentially concretized, however all such changes must be aimed at instantiation of  UHE, i.e.
+UHE can be substituted only to a term or to other UHE, but not to a variable.
+
+Let's consider an example of the technique application in building a logical inference for the following formula.
 \begin{equation*}
@@
 
-At the first step an answer $\{x \rightarrow h_1\}$ to the first question exists, $x$ is an unconfined variable and $h_1$ is an unspecified Herbrand element.
-After this step the atom $A(h_1)$ is added to the base.
-At the second step an answer $\{x \rightarrow h_2\}$ to the second question exists and $h_1$ is concretized to $f(h_2)$, after this step $B(f(h_2))$ is added to the base.
+At the first step there is an answer $\{x \rightarrow h_1\}$ to the first questions, $x$ is an unconfined variable and $h_1$ is an unspecified Herbrand element.
+After this step atom $A(h_1)$ is added to the base.
+At the second step there is an answer $\{x \rightarrow h_2\}$ to the second question and $h_1$ is instantiated to $f(h_2)$, after this step $B(f(h_2))$ is added to the base.
 Finally, at the third step the trivial (empty) answer exists to the third question and $h_2$ is concretized to $a$.
 
-For efficiency increase in this strategy, we are using two constraints.
+To improve the search efficiency of the strategy, we use two constraints.
 
@@
 %---
-Two UHEs are \emph{similar}, if they appear as the result of a substitution for the same unconfined variable.
-This situation appears if for a question with unconfined variables a sequence of answers were constructed, such that on every step of the sequence new UHE created as a substitution.
-A constraint for the similar UHE--concretizations excludes such situations, when UHE is concretized on a step, where it is not necessary, and, consequently, an opportunity to concretize it appears on other step.
-For every UHE a reference to the unconfined variable is stored, which the UHE was selected as a substitution for.
-The set of terms, which concretized a unconfined variable through its UHE, is stored in the corresponding question.
+Two UHEs are \emph{similar}, if they were generated as the result of a substitution for the same unconfined variable.
+This situation appears if for a question with unconfined variables a sequence of a similar answers were constructed, such that on every step of the sequence new UHE created as a substitution.
+The constraint for the similar UHE--concretizations excludes the possibilities of generation of new UHE when it is not necessary.
+For every UHE a reference to its unconfined variable (that resulted to the USE production) is stored.
+The set of terms, which concretized an unconfined variable through its UHE, is stored in questions.
 The engine allows tracking the UHE concretization.
-If a predefined limit of the concretizations is exceeded then all further concretizations are denied.
-The limit is chosen by user using additional information about the problem under investigation, or it is taken as $1$, and increased after an unsuccessfull logical inference, then the inference attempt is repeated.
+If a predefined limit of the concretizations is exceeded then all further concretizations rejected.
+The limit is chosen by user according to a specific and features of the problem under investigation, or it is denoted as $1$, and increased after an unsuccessfull logical inference, then the inference attempt is repeated.
 
@@
 
-For a question with unconfined variables, one can apply all available answers, i.e.
-with using all elements of Herbrand universum.
-It is known, that this is not technically possible because of unlimited Herbrand universum.
-So, the technique of lazy computations is used.
-In ideal case, all answers for question is used simultaneously, but actually only the required questions are used.
-For example, if in a some formula $\forall x: True  - \exists A(x)$, and $H^{\infty}= \{a, f(a), f(f(a)), ...\}$ then usage of all available questions simultaneously leads to addition of $\{A(a), A(f(a)), A(f(f(a))), ...\}$ elements (amount is unlimited) to the base.
-Instead, this set is associated with the element $A(h)$, where $h$ is UHE, and this element generate required elements of $H^{\infty}$ set in further.
-
-In such case, one can find a similarity with lazy concretization strategy without constraints.
-The advantage of the approach with storing expressions, which contain UHE, is that the expression with UHE is always stored, i.e.
-it is not possible to lose it by applying some substitutions.
+Another UHE--approach to process unconfined variables is based on storing in an UHE original type condition that contained the variable.
+Later it could be used as a constrain for possible substitution value from $H^{\infty}$ acceptable to form an answer.
+This is a form of lazy computation of other form as compared to lazy concretizing, but instead of sequential concretizing the USE instantiates by a term from a set $H^{\infty}$.
+The installation is stored in UHE.
+If the UHE can be instantiated in a new value from $H^{\infty}$, it does, and the value also stored in UHE.
+So we can count values one by one and prohibit repeating.
 
@@
 %-----------------k,m,-condition-----------------
-\subsection{Strategy ``k,m-condition"}
+\subsection{Strategy of ``$\boldsymbol{k,m}$-constraint"}
 An answer to a question is accepted if at some point of the following $k$ steps of its inference a certain condition becomes true at least $m$ times. This strategy is used in the following cases.
 
-\subsubsection{$k,m$--refutation.} An answer to a question with disjunctive branching is accepted if in next $k$ steps at least $m$ generated bases were refuted, otherwise the process restarted with a new answer.
+\subsubsection{$\boldsymbol{k,m}$--refutation.} An answer to a question with disjunctive branching is accepted if in next $k$ steps at least $m$ generated bases were refuted, otherwise the process restarted with a new answer.
 This strategy allows us to constrain highly spawning inferences by means of control answering questions with disjunctive branching.
 
-\subsubsection{$k$--unrefutation.}
+\subsubsection{$\boldsymbol{k}$--unrefutation.}
 For a question with disjunctive branching an answer is accepted if during next $k$ steps no refutation for the base subformula was achieved, using the questions without disjunctive branching.
@@
 If no refutation can be constructed using this strategy, then answering to the questions allowed.
-The parameter $m$ is dropped in that case.
 
-\subsubsection{$k,m$--concretization.}
+\subsubsection{$\boldsymbol{k,m}$--concretization.}
 An answer for a question is accepted if during next $k$ steps at least $m$ UHE are concretized.
@@
 After an UHE is concretized, the associations are released.
-So as soon we process the UHE as soon we will take advantage of additional memory.
+As soon we instantiate the UHE as soon we will take advantage of additional memory.
 

### 2015-09-04 paper.tex
Corrected Parallel strategies section.

@@
 %Parallel strategy
-\subsection{Parallel strategy}
+\subsection{Strategy of parallel inference}
 
@@
 
-\subsubsection{First parallel strategy.}
-If a question has disjunctive branching then, after we have answered that question, the formula will splits and transforms to the formula with number of bases equals to the number of answered question consequents, i.e.
-in general case the number of the base subformulas is increasing on every step of logical inference.
-In addition, the original formalization of a problem in PCF language can contain more than one base subformula.
+\subsubsection{Independent asynchronous refutation of bases.}
+If a question has disjunctive branching then, after we have answered that question, the formula will splits and transforms to the formula with a set of bases according to the set of answered question consequents, i.e.
+in general case the number of the base subformulas is increasing in every step of logical inference.
+In addition, the original formalization of a problem in PCF language can contain more than one base subformulas.
 In order to refute the original PCF formula every base need to be refuted.
-The refutation of the base subformulas can be processed in parallel, i.e., the refutation of each base is processed in an independent thread.
+The refutation of the base subformulas can be processed in parallel in an independent thread.
 This is possible if the bases are independent from each other, i.e.
@@
 Therefore, this strategy can be applied in case if the lazy concretization strategy is not used.
-According to the above described features of PCF we say, that PCF have property called ``OR--parallelism''.
+According to features of PCF, the PCF have property called ``OR--parallelism''.
 
-\subsubsection{Second parallel strategy.}
-In addition to the OR--parallelism, parallel schemes of algorithms for searching answers for each questions can be constructed.
+\subsubsection{Parallel answer search.}
+In addition to the OR--parallelism, parallel schemes of algorithms to search answers for each questions can be constructed.
 Since the questions do not share variables, as in previous strategy, the answer search procedure can be performed also in an individual computation thread.
@@
 
-\subsubsection{Recomendation.}
-One recommendation for an implementation of the described algorithms on cluster computing systems can be given: there is a problem to divide tasks between computing nodes depending on connection speed between the nodes.
-For example, implementation of the first strategy must bind process to the computing node.
-Binding the process to the computing node in case of second strategy is possible with increasing the conjunct of question and increasing the set of substitutions corresponding to the every atom of question.
-Otherwise, communication expenses can override the time of computation and will degrade the overall result.
+Scheduling threads and process on a cluster nodes must be carried out as usual according to the cluster parameters related to CPU performance and network throughput.
+For example, implementation of base formula refutation procedures should be bound to a process on a computing node.
+The same is not true for the second strategy, as is possible to degrade overall inference search efficiency by communication delays.
+So the refutation process in PCF having no unconfined variables are naturally scaled among cluster nodes and all the processes run independently on each other.
 
-\subsection{Memory manager}
-Allocation of memory for fixed size records is implemented based on standard approach, which is similar to \cite{gmemory}.
+\subsection{Memory management.}
+Allocation of memory for fixed size records is implemented by means of standard approach, which is similar to \cite{gmemory}.
 Array of structures is allocated with memory manager of operation system and then converted to freelists.
 New structures are allocated from the freelists, and released ones are returned back to the freelists.
-If a list is exhausted a new array is allocated from heap memory.
+If a list is exhausted a new array is allocated from heap memory if any.
 
 In the prover for each variant of records, own array is allocated and own freelist constructed.
-The garbage collection is synchronized with garbage collector of the D programming language runtime engine.
+The garbage collection is synchronized with garbage collector of the D programming language run time engine.
 

### 2015-09-04 paper.tex
Conclusion has been revised.

@@
 
-\subsection{Memory management.}
+\subsection{Memory management}
 Allocation of memory for fixed size records is implemented by means of standard approach, which is similar to \cite{gmemory}.
@@
 \section{Experiments}
-Testing the developed prover was performed on problems form TPTP \cite{tptp} library version 5.4.0 (Thousands of Problems for Theorem Provers, www.tptp.org).
-This library is the de facto standard for testing the ATP provers.
+Testing the developed prover is being performed on problems form TPTP \cite{tptp} library version 5.4.0 (Thousands of Problems for Theorem Provers, www.tptp.org).
+This library is the de facto standard for testing ATP provers.
 The type of problems that were chosen are the problems of FOF without equality.
-Total number of problems is 1221 with rating from 0.0 to 1.0.
+Total number of problems is 1221 with rating from 0.0 (very easy) to 1.0 (unresolved).
 Total number of solved problems is 780.
@@
 
-Prover shown good results for problems, whose formalization in the PCF language contains quite large conjuncts.
+Prover shown good results for problems, whose formalization in the PCF language contains large conjuncts (type conditions).
 
@@
 \section*{Conclusion}
-In this paper, we presented the results of the PCF calculus investigation in the context of the automatic theorem proving.
+In this paper, we presented the results of the positively constructed formulas (PCF) calculus investigation in the context of the automatic theorem proving.
 The main result is a prover program implemented based on the PCF calculus.
 Several logical inference strategies were devised, adopted from other papers and implemented in this prover: memory sharing, lazy concretization, $k,m$--condition, parallel strategies.
-A special global data structure (Proof state tree) has been developed.
+A special global data structure (proof state tree) has been developed.
 Most of the strategies are implemented on the base of the structure.

### 2015-09-05 paper.tex
Finished the general review.

@@
 % first the title is needed
-\title{Software for Automated Theorem Proving Based on the Calculus of Positively Constructed Formulas\thanks{Ok}}
+\title{Software for Automated Theorem Proving Based on the Calculus of Positively Constructed Formulas\thanks{\textbf{Ok .....................!!!!!!!!!!!!!!!!!!..................}}}
 % a short form should be given in case it is too long for the running head
@@
 Let $X = \{x_1,\ldots,x_k\}$ be a set of variables, $A = \{A_1,\ldots,A_m\}$ be a set of atomic formulas, and $F = \{F_1,\ldots,F_n\}$ be a set of FOFs.
-Then the following formulas $((\forall x_1) \ldots (\forall x_k) (A_1 \& \ldots \& A_m \rightarrow (F_1 \vee \ldots \vee F_n)))$ and $((\exists x_1) \ldots (\exists x_k) (A_1 \& \ldots \& A_m \& (F_1 \& \ldots \& F_n)))$ are denoted as  $\forall_XA\colon F$ and $\exists_XA\colon F$ respectively, keeping in mind that the $\forall$--quantifier corresponds to $\rightarrow F^{\vee}$, where $F^{\vee}$ means disjunction of all formulas from $F$, and $\exists$--quantifier corresponds to $\& F^{\&}$, where $F^{\&}$ means conjunctions of all formulas from $F$.
+Then the following formulas $((\forall x_1) \ldots (\forall x_k) (A_1 \& \ldots \& A_m \rightarrow (F_1 \vee \ldots \vee F_n)))$ and $((\exists x_1) \ldots (\exists x_k) (A_1 \&$ $\ldots \& A_m \& (F_1 \& \ldots \& F_n)))$ are denoted as  $\forall_XA\colon F$ and $\exists_XA\colon F$ respectively, keeping in mind that the $\forall$--quantifier corresponds to $\rightarrow F^{\vee}$, where $F^{\vee}$ means disjunction of all formulas from $F$, and $\exists$--quantifier corresponds to $\& F^{\&}$, where $F^{\&}$ means conjunctions of all formulas from $F$.
 
@@
 
-Originally, the term ``type quantifier'' was introduced by N.~Bourbaki \cite{Bourbaki} as part of notation for formalization of mathematics.
-But type quantifiers are stable in languages of another applied fields.
+\textbf{Originally, the term ``type quantifier'' was introduced by N.~Bourbaki \cite{Bourbaki} as part of notation for formalization of mathematics.
+But type quantifiers are stable in languages of another applied fields.}
 
@@
 \label{splitting}
-Let $B = \exists_XA\colon\Psi$, and $Q = \forall_YD\colon\Upsilon$, where $\Upsilon = \{\exists_{Z_1}C_1\colon\Gamma_1,\ldots,\exists_{Z_n}C_n\colon\Gamma_n\}$ then $split(B,Q) = \{\exists_{X \cup {Z_1}'} A \cup {C_1}'\colon\Psi \cup {\Gamma_1}',\ldots,\exists_{X \cup {Z_n}'} A \cup {C_n}'\colon\Psi \cup {\Gamma_n}'\}$, where $'$ is a variable renaming operator.
+Let $B = \exists_XA\colon\Psi$, and $Q = \forall_YD\colon\Upsilon$, where $\Upsilon = \{\exists_{Z_1}C_1\colon\Gamma_1,\ldots, \exists_{Z_n}C_n\colon\Gamma_n\}$ then $split(B,Q) = \{\exists_{X \cup {Z_1}'} A \cup {C_1}'\colon\Psi \cup {\Gamma_1}',\ldots,$ $\exists_{X \cup {Z_n}'} A \cup {C_n}'\colon\Psi \cup {\Gamma_n}'\}$, where $'$ is a variable renaming operator.
 We say that $B$ is split by $Q$.
@@
 \section{Experiments}
-Testing the developed prover is being performed on problems form TPTP \cite{tptp} library version 5.4.0 (Thousands of Problems for Theorem Provers, www.tptp.org).
+Testing the prover under development is being performed on problems form TPTP \cite{tptp} library version 5.4.0 (Thousands of Problems for Theorem Provers, www.tptp.org).
 This library is the de facto standard for testing ATP provers.
@@
 The maximal rating of solved problem is 0.62.
-The best results among domains presented in the Table~1.
+The best results among domains presented in Table~\ref{tbl:1}.
 
 \begin{table}
-\caption{Domains}
+\caption{Domains}\label{tbl:1}
 \begin{tabular}{|p{0.3\linewidth}|p{0.4\linewidth}|p{0.1\linewidth}|}
@@
 
-10 hardest problems solved by our prover is presented in the Table~2.
+Ten hardest problems solved by our prover is presented in Table~\ref{tbl:2}.
 \begin{table}
-\caption{Hardest problems solved by our prover}
+\caption{Hardest problems solved by our prover}\label{tbl:2}
 \begin{tabular}{|p{0.2\linewidth}|p{0.2\linewidth}|p{0.2\linewidth}|p{0.2\linewidth}|}
@@
 
-Prover shown good results for problems, whose formalization in the PCF language contains large conjuncts (type conditions).
+Prover showed good performace for problems, whose formalization in the PCF language contains large conjuncts (type conditions).
 
@@
 
-The abstract syntax tree is passed to translation component that imports the tree, translates the predicate formula or CNF, performs some reduction of subformulas by applying a set of rules, converts the formulas into the PCF language, performs additional reduction of the PCFs, renames the quantifier variables, and, at the last step, outputs the conversion result as textual representation of the problem in the language of the prover.
+The abstract syntax tree is passed to translation component that imports the tree, translates the FOF or CNF, performs some reduction of subformulas by applying a set of rules, converts the formulas into the PCF language.
+Translator also performs additional trivial reduction of the PCFs, renames the quantifier variables, and, at the last step, outputs the conversion result as textual representation of the problem in the language of the prover.
 
-\subsection{Example of Formula}
+\subsection{Example of formula}
 Problem SYN380+1.p from TPTP library in our prover input language has a following form:
@@
 
-The results of solving this problem by our prover and other provers are presented in Table~3.
+The results of solving this problem by our prover and other provers are presented in Table~\ref{tab:3}.
 
 \begin{table}
-\caption{COM003+1.p solving}
+\caption{Solution timing for COM003+1.p}\label{tab:3}
 \begin{tabular}{|p{0.2\linewidth}|p{0.35\linewidth}|}
@@
 
+In this section we list some points of the implemented strategies practical usage experience.
+
 %=============================
@@
 The decrease of the number of steps leads to a decrease of the problem solving time.
-Application of the constraint~1 is the best way of the inference time reduction in the case when a formula has questions with disjunctive branching and the constraint~2 in case if the formula contains no questions without disjunctive branching.
+\textbf{Application of the constraint~1 is the best way of the inference time reduction in the case when a formula has questions with disjunctive branching and the constraint~2 in case if the formula contains no questions without disjunctive branching}.
 
 %=============================
-\subsubsection{$k,m$--condition.}
-The efficiency of this strategy reveals in the problems with disjunctive branching.
-In all cases of acceptable application of this strategy and correctly chosen parameters, the noticeable improvement is achieved.
+\subsubsection{$\boldsymbol{k,m}$--constraint.}
+The value of this strategy reveals in the problems with disjunctive branching.
+A noticeable improvement of efficiency is achieved in all cases of acceptable application of this strategy and correctly chosen parameters.
+
 %===========================
@@
 That’s why the main characteristic of the prover productivity is not the time of execution of the program, but the ratio of program execution time in parallel mode on a given number of CNs to the program execution time on a single CN.
-The ratio is accounted for various number of CNs.
+The ratio has been evaluated for various number of CNs.
 
-Experiments were carried out on the logical problems which PCF--formalizations have necessary properties for the parallel strategies: disjunctive branching, large number of questions, large conjuncts in questions.
+Experiments were carried out on the logical problems which PCF--for\-ma\-li\-za\-ti\-ons having necessary properties for the parallel strategies utilization: disjunctive branching, large number of questions, large conjuncts in questions.
 In Fig.~\ref{fig:parallel} the results of experiments are shown.
+\textbf{In the figure ``Strategy 1'' denotes ...
+and ``Strategy 2'' denotes ...}
 

### 2015-09-06 paper.tex
Removed last questions.

@@
 % first the title is needed
-\title{Software for Automated Theorem Proving Based on the Calculus of Positively Constructed Formulas\thanks{\textbf{Ok .....................!!!!!!!!!!!!!!!!!!..................}}}
+\title{Software for Automated Theorem Proving Based on the Calculus of Positively Constructed Formulas\thanks{Research is supported by Russian Foundation of Basic Research, grant No 14-01-31449~mol\_a.}}
 % a short form should be given in case it is too long for the running head
@@
 
-\textbf{Originally, the term ``type quantifier'' was introduced by N.~Bourbaki \cite{Bourbaki} as part of notation for formalization of mathematics.
-But type quantifiers are stable in languages of another applied fields.}
+Originally, the term ``type quantifier'' was introduced by N.~Bourbaki \cite{Bourbaki} as part of notation for formalization of mathematics.
+But type quantifiers are stable in languages of another applied fields.
 
@@
   \centering
-  \includegraphics[width=0.82\linewidth]{img/DataSharing2p1.eps}
+  \includegraphics[width=0.82\linewidth]{img/DataSharing2}
   \caption{Example of PST for a base subformula refutation}
@@
 %=======================================================================
-\section{Experiments}
+\section{Testing the prover}
 Testing the prover under development is being performed on problems form TPTP \cite{tptp} library version 5.4.0 (Thousands of Problems for Theorem Provers, www.tptp.org).
@@
 The decrease of the number of steps leads to a decrease of the problem solving time.
-\textbf{Application of the constraint~1 is the best way of the inference time reduction in the case when a formula has questions with disjunctive branching and the constraint~2 in case if the formula contains no questions without disjunctive branching}.
+Application of the UHE constraint~1 (limiting depth of instantiation) is the best way of the inference time reduction in the case when a formula has questions with disjunctive branching and the constraint~2 (storing UHE expression) in case if the formula contains no questions without disjunctive branching.
 
@@
 Experiments were carried out on the logical problems which PCF--for\-ma\-li\-za\-ti\-ons having necessary properties for the parallel strategies utilization: disjunctive branching, large number of questions, large conjuncts in questions.
-In Fig.~\ref{fig:parallel} the results of experiments are shown.
-\textbf{In the figure ``Strategy 1'' denotes ...
-and ``Strategy 2'' denotes ...}
+In Fig.~\ref{fig:parallel} results of experiments are shown.
+In the figure ``Strategy 1'' denotes paralleling base subformula refutation and ``Strategy 2'' references to parallel answer search to a question.
 
@@
 \section*{Conclusion}
-In this paper, we presented the results of the positively constructed formulas (PCF) calculus investigation in the context of the automatic theorem proving.
+In this paper, we briefly presented the results of the positively constructed formulas (PCF) calculus investigation in the context of the automatic theorem proving.
 The main result is a prover program implemented based on the PCF calculus.
@@
 
+\bibitem{SNV1990} Vassilyev, S.N.: Machine Synthesis of Mathematical Theorems. The Journal of Logic programming, V.9, No.2--3, pp. 235--266 (1990)
+\bibitem{ICDS2000} Vassilyev, S.N., Zherlov, A.K., Fedunov, E.A.,
+Fedosov, B.E.: Intelligent Control of Dynamic Systems. Moscow: Fizmatlit (2000)(in Russian)
+\bibitem{Bourbaki} Bourbaki, N.: Theory of Sets. Paris, Hermann (1968)
+\bibitem{DPL1} Armstrong, J. D Programming Language. URL: http://www.digitalmars.com/d/.
 \bibitem{DPL2} Alexandrescu, A. The D Programming Language. -- Addison-Wesley Professional, 2010. -- 460p.
-
-\bibitem{DPL1} Armstrong, J. D Programming Language. URL: http://www.digitalmars.com/d/.
-
 \bibitem{Erlang1} Armstrong, J. Programming Erlang.  -- The Pragmatic Programmers, 2007. -- 519p.
-
+\bibitem{gmemory} GLIB: Memory Slices. URL: http://developer.gnome.org/glib/2.34/glib-Memory-Slices.html.
+\bibitem{tptp} Sutcliffe, G. The TPTP Problem Library and Associated Infrastructure: The FOF and CNF Parts, v3.5.0 // Journal of Automated Reasoning. V43, N4, pp.337--362.
 \bibitem{TPTPTrans} Gelder, A.V. TPTPparser utility. 2006. URL: http://users.soe.ucsc.edu/\~{}avg/TPTPparser/.
-
-\bibitem{gmemory} GLIB: Memory Slices. URL: http://developer.gnome.org/glib/2.34/glib-Memory-Slices.html.
-
 \bibitem{subtree} Graf, P. Substitution Tree Indexing // Proceedings of the 6th International Conference on Rewriting Techniques and Applications. 1995. pp.~117--131.
@@
 %\bibitem{HAR} Robinson, A., Voronkov, A. (eds.) Handbook of Automated Reasoning. -- MIT Press, Cambridge, 2001. -- 2150p.
-
-\bibitem{tptp} Sutcliffe, G. The TPTP Problem Library and Associated Infrastructure: The FOF and CNF Parts, v3.5.0 // Journal of Automated Reasoning. V43, N4, pp.337--362.
-
-\bibitem{SNV1990} Vassilyev, S.N.: Machine Synthesis of Mathematical Theorems. The Journal of Logic programming, V.9, No.2--3, pp. 235--266 (1990)
-
-\bibitem{ICDS2000} Vassilyev, S.N., Zherlov, A.K., Fedunov, E.A.,
-Fedosov, B.E.: Intelligent Control of Dynamic Systems. Moscow: Fizmatlit (2000)(in Russian)
 

### 2015-09-06 paper.tex
Spell check

@@
 
-The PCF calculus is both machine-oriented, and also human-oriented, naturally aimed at solving the problems of dynamic systems control thanks to its features such as follows: unique inference rule and simple scheme of axioms; modifiability of semantics (constructive, monotonic, temporal, etc.) and besides it is possible to construct intuitionistic inferences of some non-Horn formulas; the explicit usage of $\forall$ and $\exists$ quantifiers, the scolemization procedure is not required.
+The PCF calculus is both machine-oriented, and also human-oriented, naturally aimed at solving the problems of dynamic systems control thanks to its features such as follows: unique inference rule and simple scheme of axioms; modifyability of semantics (constructive, monotonic, temporal, etc.) and besides it is possible to construct intuitionistic inferences of some non-Horn formulas; the explicit usage of $\forall$ and $\exists$ quantifiers, the scolemization procedure is not required.
 
-However, in this paper, we will not cover the human--orientability properties of the PCF calculus and its dynamic systems control properties as well.
+However, in this paper, we will not cover the human--orientation properties of the PCF calculus and its dynamic systems control properties as well.
 This and the calculus capabilities in action planning is described in \cite{ICDS2000}.
@@
 
-Variables from $X$ are bound by correspoding quantifiers and called $\forall$--variables and $\exists$--variables, respectively.
+Variables from $X$ are bound by corresponding quantifiers and called $\forall$--variables and $\exists$--variables, respectively.
 In $\forall_XA$, a variable from $X$ that does not appear in conjunct $A$ is called {\em unconfined} variable.
@@
 Construction $\forall_XA$ and $\exists_XA$ are called positive \emph{type quantifiers} (TQ), because $A$ is a conjunction of only positive atoms and referred to as also as \emph{type condition} for $X$.
-In practise, this constructions denote phrases such as follows: ``for all $X$ satisfying $A$ there is...'' or ``there exists $X$ satisfying property $A$ such that...''; ``for all integer $x,y,z$ and $n>2$ there is $x^n + y^n \ne z^n$''.
+In practice, this constructions denote phrases such as follows: ``for all $X$ satisfying $A$ there is...'' or ``there exists $X$ satisfying property $A$ such that...''; ``for all integer $x,y,z$ and $n>2$ there is $x^n + y^n \ne z^n$''.
 
@@
 \label{answer}
-A question $\forall_YD:\Upsilon$ to a base $\exists_XA$ has an {\em answer} $\theta$ if and only if $\theta$ is a substitution $Y \rightarrow H^{\infty} \cup X$ and $D\theta \subseteq A$, where $H^{\infty}$ is Herbrand universe based on constans and function symbols that occur in corresponding base subformula.
+A question $\forall_YD:\Upsilon$ to a base $\exists_XA$ has an {\em answer} $\theta$ if and only if $\theta$ is a substitution $Y \rightarrow H^{\infty} \cup X$ and $D\theta \subseteq A$, where $H^{\infty}$ is Herbrand universe based on constant and function symbols that occur in corresponding base subformula.
 \end{definition}
@@
 The main concept of lazy instantiation is as follows.
-Instead of selecting a concrete element of a Herbrand universum for a substitution of an unconfined variable an {\em unspecified Herbrand element} (UHE) is used temporally.
+Instead of selecting a concrete element of a Herbrand universe for a substitution of an unconfined variable an {\em unspecified Herbrand element} (UHE) is used temporally.
 The UHE will be concretized further to a proper term.
-The result of the concertizing depends on the logical inference strategy.
+The result of the concretion depends on the logical inference strategy.
 UHE will be instantiated to a ground term, or in some situations, it could remain unbound.
@@
 The UHE can be concretized in a term, containing other UHEs.
-The variant of the concretizing on the each step is chosen according to inference strategies to make refutation possible, e.g.
+The variant of the concretion on the each step is chosen according to inference strategies to make refutation possible, e.g.
 by analyzing a term structure of a question.
@@
 %Для стртегии ленивой конкретизации предложено два типа ограничений
-\subsubsection{Constraint~1: Limit of concretizations number.}
+\subsubsection{Constraint~1: Limit of concretion number.}
 %Ограничение количества конкретизаций.
@@
 This situation appears if for a question with unconfined variables a sequence of a similar answers were constructed, such that on every step of the sequence new UHE created as a substitution.
-The constraint for the similar UHE--concretizations excludes the possibilities of generation of new UHE when it is not necessary.
+The constraint for the similar UHE--concretions excludes the possibilities of generation of new UHE when it is not necessary.
 For every UHE a reference to its unconfined variable (that resulted to the USE production) is stored.
 The set of terms, which concretized an unconfined variable through its UHE, is stored in questions.
-The engine allows tracking the UHE concretization.
-If a predefined limit of the concretizations is exceeded then all further concretizations rejected.
+The engine allows tracking the UHE concretion.
+If a predefined limit of the concretions is exceeded then all further concretions rejected.
 The limit is chosen by user according to a specific and features of the problem under investigation, or it is denoted as $1$, and increased after an unsuccessfull logical inference, then the inference attempt is repeated.
@@
 Later it could be used as a constrain for possible substitution value from $H^{\infty}$ acceptable to form an answer.
-This is a form of lazy computation of other form as compared to lazy concretizing, but instead of sequential concretizing the USE instantiates by a term from a set $H^{\infty}$.
+This is a form of lazy computation of other form as compared to lazy concretion, but instead of sequential concretizing the USE instantiates by a term from a set $H^{\infty}$.
 The installation is stored in UHE.
@@
 
-\subsubsection{$\boldsymbol{k,m}$--concretization.}
+\subsubsection{$\boldsymbol{k,m}$--concretion.}
 An answer for a question is accepted if during next $k$ steps at least $m$ UHE are concretized.
@@
 Bases never share $\forall$--variables because bases don't contain $\forall$--variables at all (by definitions), as they contain only ground terms.
-However, bases can share the UHEs in case of usage of the lazy concretization strategy.
-Therefore, this strategy can be applied in case if the lazy concretization strategy is not used.
+However, bases can share the UHEs in case of usage of the lazy concretion strategy.
+Therefore, this strategy can be applied in case if the lazy concretion strategy is not used.
 According to features of PCF, the PCF have property called ``OR--parallelism''.
@@
 
-%The lazy concretization (first constraint), $8$--unrefutation, $8,1$--concretization, data sharing in bases.
+%The lazy concretion (first constraint), $8$--unrefutation, $8,1$--concretion, data sharing in bases.
 
@@
 %=============================
-\subsubsection{Lazy concretization.}
-Lazy concretization strategy allows one in most cases to find a substitution for an unconfined variable in acceptable time as compared to the sequential enumeration of Herbrand universum.
+\subsubsection{Lazy concretion.}
+Lazy concretion strategy allows one in most cases to find a substitution for an unconfined variable in acceptable time as compared to the sequential enumeration of Herbrand universum.
 The constraints are the most valuable part of this strategy, which allow to decrease the number of steps of logical inference in comparison to the unconstrained strategy.
@@
 The main result is a prover program implemented based on the PCF calculus.
-Several logical inference strategies were devised, adopted from other papers and implemented in this prover: memory sharing, lazy concretization, $k,m$--condition, parallel strategies.
+Several logical inference strategies were devised, adopted from other papers and implemented in this prover: memory sharing, lazy concretion, $k,m$--condition, parallel strategies.
 A special global data structure (proof state tree) has been developed.

### 2015-09-06 paper.tex
Spell check continued. Release.

@@
 
-Prover showed good performace for problems, whose formalization in the PCF language contains large conjuncts (type conditions).
+Prover showed good performance for problems, whose formalization in the PCF language contains large conjuncts (type conditions).
 
@@
 
-The abstract syntax tree is passed to translation component that imports the tree, translates the FOF or CNF, performs some reduction of subformulas by applying a set of rules, converts the formulas into the PCF language.
+The abstract syntax tree is passed to translation component that imports the tree, translates the FOF or CNF (conjunctive normal form), performs some reduction of subformulas by applying a set of rules, converts the formulas into the PCF language.
 Translator also performs additional trivial reduction of the PCFs, renames the quantifier variables, and, at the last step, outputs the conversion result as textual representation of the problem in the language of the prover.
@@
 \subsubsection{Lazy concretion.}
-Lazy concretion strategy allows one in most cases to find a substitution for an unconfined variable in acceptable time as compared to the sequential enumeration of Herbrand universum.
+Lazy concretion strategy allows one in most cases to find a substitution for an unconfined variable in acceptable time as compared to the sequential enumeration of Herbrand universe.
 The constraints are the most valuable part of this strategy, which allow to decrease the number of steps of logical inference in comparison to the unconstrained strategy.
@@
 %===========================
-\subsubsection{Parallel stratgies.}
-The important quality of parallel schemes of algorithms is their scalability, i.e., the performance improvement respect to increase of the number of computational nodes (CNs).
+\subsubsection{Parallel strategies.}
+The important quality of parallel schemes of algorithms is their scaleability, i.e., the performance improvement respect to increase of the number of computational nodes (CNs).
 That’s why the main characteristic of the prover productivity is not the time of execution of the program, but the ratio of program execution time in parallel mode on a given number of CNs to the program execution time on a single CN.
